---
title: "Democratic Effort Power Analyses"
format:
  html:
    theme: paper
    code-tools: true
    code-fold: true
    toc: true
    toc-fold: false
    toc-location: right
    toc-depth: 3
    toc-title: "Contents"
embed-resources: true
warning: false
---

```{r}
#| include: false
# set error messages to English
Sys.setenv(LANG = "en")

knitr::opts_chunk$set(
  fig.align = "center", 
  fig.retina = 3,
  fig.width = 6, 
  fig.height = (6 * 0.618),
  out.width = "90%", 
  collapse = FALSE
)

options(
  digits = 4,
  width = 300,
  dplyr.summarise.inform = FALSE)
```

```{r}
#| include: false
#| warning: true
#| message: true
if (!require("groundhog", quietly = TRUE)) {
  install.packages("groundhog")
}
library("groundhog")

# use groundhog to install and load packages
pkgs <- c("here",         # System path management
          "tidyverse",    # ggplot, dplyr, %>%, and friends
          "DeclareDesign",    # Tools to plan, implement, analyze empirical research
          "afex",         # Analysis of factorial experiments
          "modelsummary", # Data and model summaries with tables and plots
          "tinytable",    # Lightweight package to create tables
          "marginaleffects",      # Compute marginal means, slopes, predictions, etc.
          "patchwork",    # Combine multiple ggplots
          "showtext",    # Additional fonts for plots etc
          "tictoc"        # Timing functions
          )

groundhog.library(pkgs, "2026-01-01")
```

```{r}
#| include: false

showtext_auto()
showtext_opts(dpi = 600)

font_add_google("Public Sans", "Public Sans")
font_add_google("Public Sans", "Public Sans Light", regular.wt = 300)
font_add_google("Public Sans", "Public Sans Semibold", regular.wt = 600)

# Custom ggplot theme
theme_nice <- function(base_size = 14) {
  # General text family and base size
  theme_minimal(base_family = "Public Sans", base_size = base_size) +
    theme(
      # Remove minor grid lines for cleaner look
      panel.grid.minor = element_blank(),
      # Customize major grid lines color
      panel.grid.major = element_blank(),
      # Plot title: bold and custom relative size
      plot.title = element_text(family = "Public Sans", face = "bold", size = rel(1.1), margin = margin(b = 10)),
      # Subtitle and caption: lighter weight 
      plot.subtitle = element_text(family = "Public Sans Light", face = "plain"),
      plot.caption = element_text(family = "Public Sans Light", face = "plain"),
      # Axis text: Relative size
      axis.text = element_text(size = rel(0.9)),
      # Axis titles: Semibold and custom relative size
      axis.title = element_text(family = "Public Sans Semibold", size = rel(0.9)),
      axis.title.x = element_text(hjust = 0.5),
      axis.title.y = element_text(hjust = 0.5),
      # Axis lines
      axis.line = element_line(color = "grey25", linewidth = 0.8),
      # Facet labels: bold text on grey background
      strip.text = element_text(
        family = "Public Sans Semibold", face = "plain",
        size = rel(1),
        hjust = 0.5
      ),
      strip.background = element_rect(fill = "grey95", color = "grey25", linewidth = 0.8), 
      # Legend styling: compact and top-positioned
      legend.title = element_text(family = "Public Sans Semibold", size = rel(0.8)),
      legend.text = element_text(size = rel(0.8)),
      legend.position = "top",
      legend.justification = "center",
      legend.title.position = "left",
      legend.margin = margin(l = 0, t = 0)
    )
}


theme_set(theme_nice())
update_geom_defaults("text", list(family = "Public Sans"))
update_geom_defaults("label", list(family = "Public Sans"))
```

# Democratic Effort Task Power Analyses 

We want to simulate data to calculate the power to reliably detect between-subject differences of Cohens D = 0.25 in (k_ingroup - k_outgroup) between the correction treatment vs. control conditions. 

## 1. Power to detect treatment differences in k_in_out given the effect size of interest of Cohens D = 0.25

### Setup 

Declare the model
```{r}
# Parameters
N <- 500  # Total sample size (250 per group)
cohens_d <- -0.25 # Effect size of interest (negative = lower in-group favouritism)
sd_outcome <- 0.4 # SD from Study 1
mean_control <- 0.2  # Mean from Study 1 (control condition baseline)
effect_size <- cohens_d * sd_outcome  # Raw effect = -0.10

# Declare model: between-subjects design with potential outcomes
model <- declare_model(
  N = N,
  # Baseline outcome under control condition
  Y_Z_0 = rnorm(N, mean = mean_control, sd = sd_outcome),
  # Potential outcome under treatment: shift by raw effect
  Y_Z_1 = Y_Z_0 + effect_size
)
```

Declare the inquiry (estimate)
```{r}
inquiry <- declare_inquiry(
  ATE = mean(Y_Z_1 - Y_Z_0)
)
```

Declare the data strategy
```{r}
# Assignment: balanced random assignment to treatment (Z=1) or control (Z=0)
assignment <- declare_assignment(
  Z = complete_ra(N = N, conditions = c(0, 1)),
  Y = reveal_outcomes(Y ~ Z)
)
```

Declare the estimator (paired t-test)
```{r}
estimator <- declare_estimator(
  Y ~ Z,
  model = lm_robust,
  inquiry = "ATE",
  label = "OLS"
)
```

Declare the diagnosands
```{r}
diagnosands <- declare_diagnosands(
  # Bias and accuracy
  bias            = mean(estimate - estimand),
  mean_estimate   = mean(estimate),
  mean_estimand   = mean(estimand),
  
  # Variance and efficiency
  sd_estimate     = sd(estimate),
  rmse            = sqrt(mean((estimate - estimand)^2)),
  
  # Power and error rates
  power           = mean(p.value < 0.05),
  type_s_rate     = mean(sign(estimate) != sign(estimand)),
  
  # Confidence intervals
  coverage        = mean(conf.low <= estimand & estimand <= conf.high),
  ci_lower        = mean(conf.low, na.rm = TRUE),
  ci_upper        = mean(conf.high, na.rm = TRUE),
  mean_ci_width   = mean(conf.high - conf.low, na.rm = TRUE),
  
  # Effect size: standardised by the population SD from Study 1
  cohens_d        = mean(estimate) / sd_outcome,
  
  # Sample size
  mean_N          = mean(N)
)
```

Put the design together
```{r}
design_between <- model + assignment + inquiry + estimator

# Test simulation at N = 500 (250 per group)
diagnosis_test <- diagnose_design(
  design_between,
  sims = 1000,
  diagnosands = diagnosands
)

diagnosis_test
```

### Simulation

```{r}
#| warning: false
#| message: false
designs_list_between <- redesign(design_between,
                                 N = seq(400, 700, by = 50))

n_sim <- 2000
cat("Running", length(designs_list_between), "designs with", n_sim, "simulations each...\n")
cat("Total simulations:", length(designs_list_between) * n_sim, "\n")
cat("N per group ranges from", 400/2, "to", 700/2, "\n")

tic("Power analysis - between subjects")
diagnosis_between <- diagnose_design(designs_list_between,
                                     sims = n_sim,
                                     diagnosands = diagnosands)
toc()
```

### Visualisation

```{r}
fig_power_data <- diagnosis_between$diagnosands_df |>
  mutate(n_per_group = N / 2)

fig_power <- fig_power_data |>
  ggplot(aes(x = n_per_group, y = power)) +
  geom_line(color = "firebrick3",
            linewidth = 1) +
  geom_point(color = "firebrick3",
             size = 2) +
  geom_hline(yintercept = 0.80,
             linetype = "dashed",
             color = "black",
             linewidth = 0.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0.5, 1),
                     breaks = seq(0, 1, by = 0.1)) +
  scale_x_continuous(breaks = seq(200, 350, by = 25)) +
  labs(title = "Power to detect Cohen's d = -0.25",
       subtitle = "Between-subjects design, two balanced groups",
       x = "Sample size (n per group)",
       y = "Power",
       caption = "Dashed line = 80% power threshold. Based on 2000 simulations per design.") +
  theme_nice()

fig_power
```

```{r}
between_sim_table <- diagnosis_between$diagnosands_df |>
  transmute(
    "N (total)"     = N,
    "N (per group)" = N / 2,
    "Cohen's d"     = round(cohens_d, 2),
    "Power"         = round(power, 2),
    "SE (Power)"    = round(`se(power)`, 3),
    "Type S Rate"   = round(type_s_rate, 3),
    "Coverage"      = round(coverage, 2)
  ) |>
  tt( notes = "Power diagnosis based on 2000 simulations per design. Bootstrapped standard errors for power estimates. Type S rate = proportion of simulations where effect sign is incorrect.")

between_sim_table
```


