---
title: "Democratic Effort Study 2: Descriptives"
format:
  html:
    theme: paper
    code-tools: true
    code-fold: true
    toc: true
    toc-fold: false
    toc-location: right
    toc-depth: 3
    toc-title: "Contents"
embed-resources: true
warning: false
---

```{r}
#| include: false
# set error messages to English
Sys.setenv(LANG = "en")

knitr::opts_chunk$set(
  fig.align = "center", 
  fig.retina = 3,
  fig.width = 6, 
  fig.height = (6 * 0.618),
  out.width = "90%", 
  collapse = FALSE
)

options(
  digits = 4,
  width = 300,
  dplyr.summarise.inform = FALSE)
```

```{r}
#| include: false
#| warning: true
#| message: true
if (!require("groundhog", quietly = TRUE)) {
  install.packages("groundhog")
}
library("groundhog")

# use groundhog to install and load packages
pkgs <- c("here",         # System path management
          "datawizard",   # Data cleaning & standardization
          "tidyverse",    # ggplot, dplyr, and friends
          "modelsummary", # Data and model summaries with tables and plots
          "tinytable",    # Lightweight package to create tables
          "effectsize",   # Effect sizes for common tests
          "GGally",       # ggpairs and other ggplot2 friends
          "ggdist",       # Additional plot elements
          "corrplot",     # Visualise correlations
          "ggrain",       # For fancy raincloud plots
          "ggsignif",     # Add significance to plots
          "patchwork",    # Combine plots
          "showtext"      # Additional fonts for plots etc
          )


groundhog.library(pkgs, "2026-01-01")
```

```{r}
#| include: false

showtext_auto()
showtext_opts(dpi = 600)

font_add_google("Public Sans", "Public Sans")
font_add_google("Public Sans", "Public Sans Light", regular.wt = 300)
font_add_google("Public Sans", "Public Sans Semibold", regular.wt = 600)

# Custom ggplot theme
theme_nice <- function(base_size = 14) {
  # General text family and base size
  theme_minimal(base_family = "Public Sans", base_size = base_size) +
    theme(
      # Remove minor grid lines for cleaner look
      panel.grid.minor = element_blank(),
      # Customize major grid lines color
      panel.grid.major = element_blank(),
      # Plot title: bold and custom relative size
      plot.title = element_text(family = "Public Sans", face = "bold", size = rel(1.1), margin = margin(b = 10)),
      # Subtitle and caption: lighter weight 
      plot.subtitle = element_text(family = "Public Sans Light", face = "plain"),
      plot.caption = element_text(family = "Public Sans Light", face = "plain"),
      # Axis text: Relative size
      axis.text = element_text(size = rel(0.9)),
      # Axis titles: Semibold and custom relative size
      axis.title = element_text(family = "Public Sans Semibold", size = rel(0.9)),
      axis.title.x = element_text(hjust = 0.5),
      axis.title.y = element_text(hjust = 0.5),
      # Axis lines
      axis.line = element_line(color = "grey25", linewidth = 0.8),
      # Facet labels: bold text on grey background
      strip.text = element_text(
        family = "Public Sans Semibold", face = "plain",
        size = rel(1),
        hjust = 0.5
      ),
      strip.background = element_rect(fill = "grey95", color = "grey25", linewidth = 0.8), 
      # Legend styling: compact and top-positioned
      legend.title = element_text(family = "Public Sans Semibold", size = rel(0.8)),
      legend.text = element_text(size = rel(0.8)),
      legend.position = "top",
      legend.justification = "center",
      legend.title.position = "left",
      legend.margin = margin(l = 0, t = 0)
    )
}

theme_set(theme_nice())
update_geom_defaults("text", list(family = "Public Sans"))
update_geom_defaults("label", list(family = "Public Sans"))

agent_limits <- c("VOTERURAL", "VOTEURBAN")
agent_colors <- c("VOTERURAL" = "firebrick3", "VOTEURBAN" = "slateblue4")  
agent_fills <- c("VOTERURAL" = "firebrick3", "VOTEURBAN" = "slateblue4")   
agent_lines <- c("VOTERURAL" = "dashed", "VOTEURBAN" = "dashed")

party_colors <- c("Democrat" = "slateblue4", "Republican" = "firebrick3")
party_fills  <- c("Democrat" = "slateblue4", "Republican" = "firebrick3")

benefit_limits_lab <- c("In-Group",  "Out-Group")
benefit_colors_lab <- c("In-Group" = "#009988", "Out-Group" = "#EE7733") 
benefit_fills_lab <- c("In-Group" = "#009988", "Out-Group" = "#EE7733")
benefit_lines_lab <- c("In-Group" = "dashed", "Out-Group" = "dashed")
```

```{r}
#| include: false
data_dir <- here("01_data", "final") 
table_dir <- here("02_output", "tables", "descriptives") 
fig_dir <- here("02_output", "figures", "descriptives") 

dir.create(table_dir, recursive = TRUE, showWarnings = TRUE)
dir.create(fig_dir, recursive = TRUE, showWarnings = TRUE)
```

```{r}
load(here(data_dir, "demeffort_survey_data.RData")) 
load(here(data_dir, "demeffort_combined_data.RData"))
```

## Participant descriptives

```{r}
descriptive_table <- survey_data |>
  mutate(across(where(is.factor), droplevels)) |>
  select(
    Age = age,
    Gender = gender,
    Party = party,
    "Rural / Urban" = rururb01,
    Education = education,
    "Primary Ethnicity" = ethnicity01,
    Income = income
  ) |>
  datasummary_skim(
    type = "all",
    fun_numeric = list(
      Mean   = Mean,
      SD     = SD,
      Min    = Min,
      Median = Median,
      Max    = Max
    )
  ) |>
  style_tt(j = 1, align = "l") |>
  style_tt(j = 2:3, align = "c")

descriptive_table |> 
  save_tt(here(table_dir, "sample_descriptives.docx"), overwrite = T)

descriptive_table
```

```{r}
options(modelsummary_balance_fn = list(Mean = Mean, SD = SD, Min = Min, Median = Median, Max = Max))

descriptive_party_table <- survey_data |>
  mutate(across(where(is.factor), droplevels)) |>
  select(
    Age = age,
    Gender = gender,
    "Rural / Urban" = rururb01,
    Education = education,
    "Primary Ethnicity" = ethnicity01,
    Income = income,
    Party = party
  ) |>
  datasummary_balance(~Party, data = _) |>
  style_tt(j = 1, align = "l") |>
  style_tt(j = 2:3, align = "c")

descriptive_party_table |>
  save_tt(here(table_dir, "by_party_sample_descriptives.docx"), overwrite = TRUE)
descriptive_party_table
```

```{r}
descriptive_exp_table <- survey_data |>  
  select(
    Condition = condition,
    Age = age,
    Gender = gender,
    "Rural / Urban" = rururb01,
    Education = education,
    "Primary Ethnicity" = ethnicity01,
    Income = income,
    Party = party
  ) |> 
  droplevels() |> 
  datasummary_balance(~Condition, data = _, dinm = FALSE) |> 
  style_tt(j = 1, align = "l") |> 
  style_tt(j = 2:3, align = "c")

descriptive_exp_table |> 
  save_tt(here(table_dir, "by_condition_sample_descriptives.docx"), overwrite = T)

descriptive_exp_table
```

## Variable Descriptives

```{r}
survey_data |>  
  mutate(Duration = as.numeric(timer_total) / 60) |> 
  select(Duration) |> 
  datasummary_skim() |> 
  style_tt(j = 1, align = "l") |> 
  style_tt(j = 2:3, align = "c")
```

```{r}
variable_table <- survey_data |>
  select(
    "K In-group"               = k_ingroup,
    "K Out-group"              = k_outgroup,
    "K Rural"                  = k_rural,
    "K Urban"                  = k_urban,
    "K Out-group − In-group"   = k_outgroup_ingroup,
    "K Mean"                   = k_mean,
    "β (Inverse Temperature)"  = beta,
    "Deviation"                = deviation,
    "Affective Polarization"   = affpol_weighted,
    "Party Importance"         = party_importance_raw,
    "Democratic Support"       = democratic_support_raw,
    "Free & Fair Elections"    = freefair_raw
  ) |>
  datasummary_skim(
    type = "numeric",
    fun_numeric = list(
      Mean   = Mean,
      SD     = SD,
      Min    = Min,
      Median = Median,
      Max    = Max
    )
  ) |>
  style_tt(j = 1, align = "l") |>
  style_tt(j = 2:3, align = "c")

variable_table |>
  save_tt(here(table_dir, "variable_descriptives.docx"), overwrite = TRUE)
variable_table
```

```{r}
variable_party_table <- survey_data |>  
  select(
    Party                      = party_preference,
    "K In-group"               = k_ingroup,
    "K Out-group"              = k_outgroup,
    "K Rural"                  = k_rural,
    "K Urban"                  = k_urban,
    "K Out-group − In-group"   = k_outgroup_ingroup,
    "K Mean"                   = k_mean,
    "β (Inverse Temperature)"  = beta,
    Deviation                  = deviation,
    "Affective Polarization"   = affpol_weighted,
    "Party Importance"         = party_importance_raw,
    "Democratic Support"       = democratic_support_raw,
    "Free & Fair Elections"    = freefair_raw
  ) |> 
  droplevels() |> 
  datasummary_balance(~Party, data = _) |> 
  style_tt(j = 1, align = "l") |> 
  style_tt(j = 2:3, align = "c")

variable_party_table |> 
  save_tt(here(table_dir, "variable_party_descriptives.docx"), overwrite = TRUE)
variable_party_table
```

```{r}
variable_exp_table <- survey_data |>  
  select(
    Condition                  = condition,
    "K In-group"               = k_ingroup,
    "K Out-group"              = k_outgroup,
    "K Rural"                  = k_rural,
    "K Urban"                  = k_urban,
    "K Out-group − In-group"   = k_outgroup_ingroup,
    "K Mean"                   = k_mean,
    "β (Inverse Temperature)"  = beta,
    Deviation                  = deviation,
    "Affective Polarization"   = affpol_weighted,
    "Party Importance"         = party_importance_raw,
    "Democratic Support"       = democratic_support_raw,
    "Free & Fair Elections"    = freefair_raw
  ) |> 
  droplevels() |> 
  datasummary_balance(~Condition, data = _) |> 
  style_tt(j = 1, align = "l") |> 
  style_tt(j = 2:3, align = "c")

variable_exp_table |> 
  save_tt(here(table_dir, "variable_condition_descriptives.docx"), overwrite = TRUE)
variable_exp_table
```

## Balance Tables

### Benefit In-Group / Out-Group / Everyone

```{r}
benefit_balance_table <- combined_data |> 
  select(decision, 
         benefit) |> 
  mutate(
    Decision = case_match(
      decision,
      1 ~ "Work",
      0 ~ "Rest",
      NA ~ "NaN"),
    Condition = case_match(
      benefit, 
      "INGROUP" ~ "In-Group",
      "OUTGROUP" ~ "Out-Group"
    )
  ) |> 
  mutate(Condition = factor(Condition, levels = c("In-Group", "Out-Group"))) |> 
  datasummary_crosstab(Condition ~ Decision, 
                       data = _)

benefit_balance_table |> 
  save_tt(here(table_dir, "benefit_balance_table.docx"), overwrite = T)

benefit_balance_table
```

### Within Factor 2: Effort

```{r}
combined_data |> 
  mutate(
    Decision = case_match(
      decision,
      1 ~ "Work",
      0 ~ "Rest",
      NA ~ "NaN"
    ),
    Condition = factor(
      case_match(
        eff, 
        "easy65" ~ "Low Effort",
        "hard80" ~ "Mid Effort",
        "hard95" ~ "High Effort"
      ),
      levels = c("Low Effort", "Mid Effort", "High Effort")
    )
  ) |> 
  datasummary_crosstab(
    Condition ~ Decision, 
    data = _
  )
```

### Follow-Up Agent x Partisanship

```{r}
agent_balance_table <- combined_data |> 
  select(decision, 
         agent,
         party) |> 
  droplevels() |> 
  mutate(
    Decision = case_match(
      decision,
      1 ~ "Work",
      0 ~ "Rest",
      NA ~ "NaN"),
    Condition = case_match(
      agent, 
      "VOTERURAL" ~ "Rural",
      "VOTEURBAN" ~ "Urban"
    )
  ) |> 
  mutate(Condition = factor(Condition, levels = c("Urban", "Rural"))) |> 
  datasummary_crosstab(Decision ~ Condition * party, 
                       statistic = ~ Percent("col"),
                       data = _)

agent_balance_table |> 
  save_tt(here(table_dir, "agent_balance_table.docx"), overwrite = T)

agent_balance_table
```

### Follow-Up Condition x Benefit

```{r}
treatment_balance_table <- combined_data |> 
  select(decision, 
         benefit,
         Correction = condition) |> 
  droplevels() |> 
  mutate(
    Decision = case_match(
      decision,
      1 ~ "Work",
      0 ~ "Rest",
      NA ~ "NaN"),
    Benefit = case_match(
      benefit, 
      "INGROUP" ~ "In-Group",
      "OUTGROUP" ~ "Out-Group"
    )
  ) |> 
  mutate(Benefit = factor(Benefit, levels = c("In-Group", "Out-Group"))) |>  
  datasummary_crosstab(Decision ~ Correction * Benefit, 
                       statistic = ~ Percent("col"),
                       data = _)

treatment_balance_table |> 
  save_tt(here(table_dir, "treatment_balance_table.docx"), overwrite = T)

treatment_balance_table
```

## Correlations

```{r}
#| message: false
#| fig-width: 10
#| fig-height: 9
#| fig-dpi: 600
scatter_function <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_point(alpha = 0.2, size = 1, color = "#fd8d3c") +
    geom_smooth(method = "lm",
                se = TRUE,
                linewidth = 1.2,
                alpha = 0.15,
                color = "black",
                fill = "black",
                na.rm = TRUE)
}

corr_plot <- survey_data |>
  select(
    "K Out−In"  = k_outgroup_ingroup,
    "K Mean"    = k_mean,
    "Beta"      = beta,
    "Deviation" = deviation,
    "AffPol."   = affpol_weighted,
    "PartImp"   = party_importance_raw,
    "DemSupp"   = democratic_support_raw,
    "FreeFair"  = freefair_raw
  ) |>
  ggpairs(
    lower = list(continuous = scatter_function),
    upper = list(continuous = wrap("cor", stars = TRUE, size = 5)),
    diag  = list(continuous = wrap("densityDiag", alpha = 0.5,
                                   fill = "#fd8d3c", color = "#f16913")),
    progress = FALSE
  ) +
  theme_nice() +
  theme(
    axis.text  = element_text(size = 10),
    strip.text = element_text(size = 11)
  )

corr_plot

ggsave(
  here(fig_dir, "variable_correlogram.png"),
  height = 9,
  width  = 10,
  dpi    = 600
)
```


## In-Group / Out-Group Figures

### Regression Plot

```{r}
p_dat <- combined_data |> 
  filter(!is.na(decision)) |>
    mutate(
    effort_num = case_when(
      eff == "easy65" ~ 1,
      eff == "hard80" ~ 2, 
      eff == "hard95" ~ 3),
    reward_num = as.numeric(as.character(rew)),
    benefit_label = factor(benefit,
                           levels = c("INGROUP", "OUTGROUP"),
                           labels = c("In-Group", "Out-Group")))
```

```{r}
p_effort_reg <- p_dat |>
  ggplot(aes(x = effort_num, y = decision, 
             color = benefit_label, fill = benefit_label, linetype = benefit_label)) +
  stat_smooth(alpha = 0.3, linewidth = 1.4, method = "glm", method.args = list(family = "binomial")) +
  stat_smooth(alpha = 0, linewidth = 1.4, method = "glm", method.args = list(family = "binomial")) +
  scale_fill_manual(limits = benefit_limits_lab, values = benefit_fills_lab) +
  scale_color_manual(limits = benefit_limits_lab, values = benefit_colors_lab) +
  scale_linetype_manual(limits = benefit_limits_lab, values = benefit_lines_lab) +
  scale_x_continuous(
    breaks = c(1, 2, 3),
    labels = c("1", "2", "3"),
    expand = expansion(mult = c(0.2, 0.2))
  ) +
  scale_y_continuous(
    breaks = seq(0, 1.0, by = 0.25),
    limits = c(0, 1)
  ) +
  theme_nice() +
  labs(x = "Effort Level", 
       y = "Choices to Work") +
  theme(legend.position = "bottom")
```

```{r}
p_reward_reg <- p_dat |>
  ggplot(aes(x = reward_num, y = decision, 
             color = benefit_label, fill = benefit_label, linetype = benefit_label)) +
  stat_smooth(alpha = 0.3, linewidth = 1.4, method = "glm", method.args = list(family = "binomial")) +
  stat_smooth(alpha = 0, linewidth = 1.4, method = "glm", method.args = list(family = "binomial")) +
  scale_fill_manual(limits = benefit_limits_lab, values = benefit_fills_lab) +
  scale_color_manual(limits = benefit_limits_lab, values = benefit_colors_lab) +
  scale_linetype_manual(limits = benefit_limits_lab, values = benefit_lines_lab) +
  scale_x_continuous(
    breaks = c(4, 12, 20),
    labels = c("4", "12", "20"),
    expand = expansion(mult = c(0.2, 0.2))
  ) +
  scale_y_continuous(
    breaks = seq(0, 1.0, by = 0.25),
    limits = c(0, 1)
  ) +
  theme_nice() +
  labs(x = "Reward Available", 
       y = "Choices to Work") +
  theme(legend.position = "bottom")
```

```{r}
#| fig-width: 5
#| fig-height: 6
#| fig-dpi: 600
showtext_opts(dpi = 600)

fig_work_reg <- ((p_effort_reg / p_reward_reg) ) +
  plot_layout(guides = "collect") +
  plot_annotation(title = "Study 1") &
  theme(legend.title = element_blank())

fig_work_reg

ggsave(here(fig_dir, "fig_work_reg.png"),
       width = 5, height = 6, dpi = 600)
```

### K Plot

```{r}
dev_data <- survey_data |>
  filter(deviation < 0) |>
  mutate(condition = factor(condition, levels = c("Treatment", "Control")))

t_ingroup   <- t.test(k_ingroup  ~ condition, data = dev_data)
t_outgroup  <- t.test(k_outgroup ~ condition, data = dev_data)
t_ctrl      <- t.test(dev_data$k_ingroup[dev_data$condition == "Control"],
                      dev_data$k_outgroup[dev_data$condition == "Control"],
                      paired = TRUE)
t_treat     <- t.test(dev_data$k_ingroup[dev_data$condition == "Treatment"],
                      dev_data$k_outgroup[dev_data$condition == "Treatment"],
                      paired = TRUE)

d_ingroup   <- cohens_d(k_ingroup  ~ condition, data = dev_data)$Cohens_d
d_outgroup  <- cohens_d(k_outgroup ~ condition, data = dev_data)$Cohens_d
d_ctrl      <- cohens_d(dev_data$k_ingroup[dev_data$condition == "Control"],
                        dev_data$k_outgroup[dev_data$condition == "Control"],
                        paired = TRUE)$Cohens_d
d_treat     <- cohens_d(dev_data$k_ingroup[dev_data$condition == "Treatment"],
                        dev_data$k_outgroup[dev_data$condition == "Treatment"],
                        paired = TRUE)$Cohens_d

signif_df <- tibble(
  group1     = c(0.65, 1.35, 0.65, 1.65),
  group2     = c(1.65, 2.35, 1.35, 2.35),
  p.value    = c(t_ingroup$p.value, t_outgroup$p.value, t_treat$p.value, t_ctrl$p.value),
  cohens_d   = c(d_ingroup, d_outgroup, d_treat, d_ctrl),
  y_position = c(2.2, 2.45, 1.75, 1.75)
) |> 
  mutate(
    stars = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE           ~ "ns"
    ),
    label = paste0(stars, " (d = ", round(cohens_d, 2), ")")
  )

dat_k <- dev_data |>
  select(PID, condition, k_ingroup, k_outgroup) |>
  pivot_longer(
    cols      = c(k_ingroup, k_outgroup),
    names_to  = "benefit",
    values_to = "k"
  ) |>
  mutate(
    benefit_label = factor(benefit,
                           levels = c("k_ingroup", "k_outgroup"),
                           labels = c("In-Group", "Out-Group"))
  )

dat_k_sum <- dat_k |>
  summarise(
    mean_k = mean(k),
    se_k   = sd(k) / sqrt(n()),
    .by    = c(condition, benefit_label)
  )
```

```{r}
#| fig-width: 4
#| fig-height: 3
#| fig-dpi: 600

p_k <- ggplot() +
  geom_col(
    data     = dat_k_sum,
    aes(x = condition, y = mean_k, fill = benefit_label),
    position = position_dodge(width = 0.7),
    color = "black", alpha = 0.9, width = 0.6
  ) +
  geom_jitter(
    data  = dat_k,
    aes(x = condition, y = k, fill = benefit_label),
    shape = 21, color = "black",
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.7),
    alpha = 0.3, size = 0.9
  ) +
  geom_errorbar(
    data  = dat_k_sum,
    aes(x = condition, ymin = mean_k - se_k, ymax = mean_k + se_k,
        group = benefit_label),
    position = position_dodge(width = 0.7),
    width = 0.15, linewidth = 0.5
  ) +
  geom_signif(
    data       = signif_df,
    aes(xmin = group1, xmax = group2, annotations = label, y_position = y_position),
    manual     = TRUE,
    tip_length = 0,
    size       = 0.8,
    textsize   = 3.5,
    family     = "Public Sans",
    color      = "black"
  ) +
  scale_fill_manual(values = benefit_fills_lab, name = NULL) +
  scale_y_continuous(
    limits = c(0, 2.5),
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(title = "Effort Discounting (K)",
       x = NULL,
       y = "Effort Discounting (K)") +
  theme(legend.position = "none")

ggsave(here(fig_dir, "fig_k.png"), p_k, width = 4, height = 3, dpi = 600)
```

##### Collapsed across conditions

```{r}
# Data & tests (collapsed across condition)
t1_collapsed <- t.test(dev_data$k_ingroup, dev_data$k_outgroup, paired = TRUE)
d1_collapsed <- cohens_d(dev_data$k_ingroup, dev_data$k_outgroup, paired = TRUE)$Cohens_d

signif_df_collapsed <- tibble(
  group1     = c("In-Group"),
  group2     = c("Out-Group"),
  p.value    = c(t1_collapsed$p.value),
  cohens_d   = c(d1_collapsed),
  y_position = c(1.9)
) |>
  mutate(
    stars = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE           ~ "ns"
    ),
    label = paste0(stars, " (d = ", round(cohens_d, 2), ")")
  )

dat_k_collapsed <- dev_data |>
  select(PID, k_ingroup, k_outgroup) |>
  pivot_longer(
    cols      = c(k_ingroup, k_outgroup),
    names_to  = "benefit",
    values_to = "k"
  ) |>
  mutate(
    benefit_label = factor(benefit,
                           levels = c("k_ingroup", "k_outgroup"),
                           labels = c("In-Group", "Out-Group"))
  )

dat_k_sum_collapsed <- dat_k_collapsed |>
  summarise(
    mean_k = mean(k),
    se_k   = sd(k) / sqrt(n()),
    .by    = benefit_label
  )
```

```{r}
#| fig-width: 4
#| fig-height: 3
#| fig-dpi: 600
showtext.opts(dpi = 600)
p_k_collapsed <- ggplot() +
  geom_col(
    data  = dat_k_sum_collapsed,
    aes(x = benefit_label, y = mean_k, fill = benefit_label),
    width = 0.6,
    color = "black", alpha = 0.9
  ) +
  geom_jitter(
    data  = dat_k_collapsed,
    aes(x = benefit_label, y = k, fill = benefit_label),
    shape = 21, color = "black",
    width = 0.1, alpha = 0.3, size = 0.9
  ) +
  geom_errorbar(
    data  = dat_k_sum_collapsed,
    aes(x = benefit_label, ymin = mean_k - se_k, ymax = mean_k + se_k),
    width = 0.15, linewidth = 0.5
  ) +
  geom_signif(
    data       = signif_df_collapsed,
    aes(xmin = group1, xmax = group2, annotations = label, y_position = y_position),
    manual     = TRUE,
    tip_length = 0,
    size       = 0.8,
    textsize   = 3.5,
    family     = "Public Sans",
    color      = "black"
  ) +
  scale_fill_manual(values = benefit_fills_lab, name = NULL) +
  scale_y_continuous(
    limits = c(0, 2),
    breaks = seq(0, 2, by = 0.5),
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(x = NULL,
       y = "Effort Discounting (K)") +
  theme(legend.position = "none")

ggsave(here(fig_dir, "fig_k_collapsed.png"), p_k_collapsed, width = 4, height = 3, dpi = 600)
```

### Combined Regression Plot Figure

```{r}
#| fig-width: 9
#| fig-height: 4
#| fig-dpi: 600
showtext_opts(dpi = 600)

fig_reg <- wrap_plots(
    p_effort_reg, p_reward_reg, p_k_collapsed, guide_area(),
    design  = c(area(1,1), area(1,2), area(1,3), area(2,1,2,3)),
    widths  = c(1, 1, 1.5),
    heights = c(10, 1),
    guides  = "collect"
  ) +
  plot_annotation(
    title = "Study 2",
    theme = theme(plot.title = element_text(size = 18, face = "bold",
                                            family = "Public Sans"))
  ) &
  theme(legend.title    = element_blank())

fig_reg

ggsave(here(fig_dir, "fig_reg.png"),
       width = 9, height = 4, dpi = 600)
```

```{r}
#| fig-width: 9
#| fig-height: 5
#| fig-dpi: 600
showtext_opts(dpi = 600)

fig_reg_condition <- ((p_effort_reg / p_reward_reg) | p_k) +
  plot_layout(guides = "collect", widths = c(1, 1.8)) +
  plot_layout(guides = "collect") +
  plot_annotation(title = "Study 2") &
  theme(legend.title = element_blank())

fig_reg_condition

ggsave(here(fig_dir, "fig_reg_condition.png"),
       width = 9, height = 5, dpi = 600)
```

## Partisanship Figures

```{r}
signif_df_fav <- tibble(
  group1     = "Democrat",
  group2     = "Republican",
  label      = "*** (b = -0.30)",
  y_position = 3
)
```

```{r}
p_fav_partisan <- 
  survey_data |> 
  mutate(k_favoritism = standardise(k_outgroup_ingroup, center = 0)) |> 
  ggplot(
    aes(x = party, y = k_favoritism, fill = party, color = party)
  ) +
  geom_hline(
    yintercept = 0, linetype = "dashed",
    color = "grey50", linewidth = 0.4
  ) +
  geom_rain(
    alpha        = 0.4,
    rain.side    = "f1x1",
    violin.args  = list(color = "black", alpha = 0.4, width = 0.5),
    boxplot.args = list(color = "black", outlier.shape = NA, width = 0.1),
    point.args   = list(shape = 21, color = "black", size = 0.7, alpha = 0.25)
  ) +
  geom_signif(
    data        = signif_df_fav,
    aes(xmin = group1, xmax = group2,
        annotations = label, y_position = y_position),
    manual      = TRUE,
    tip_length  = 0,
    size        = 0.8,
    textsize    = 3.5,
    family      = "Public Sans",
    color       = "black",
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values  = party_fills,  name = NULL) +
  scale_color_manual(values = party_colors, name = NULL) +
  scale_y_continuous(
    breaks = seq(-3, 3, by = 1),
    expand = expansion(mult = c(0.05, 0.1))
  ) +
  labs(x = NULL, y = "In-Group Favoritism") +
  theme(legend.position = "none")

p_fav_partisan
```

```{r}
b_val = 0.27

stats_label <- paste0("*** ", "(b = ", b_val, ")")

p_scatter <- survey_data |>
  select(affpol_weighted, k_outgroup_ingroup) |>
  mutate(affpol_weighted = as.numeric(affpol_weighted)) |>
  standardise(select = c("affpol_weighted", "k_outgroup_ingroup")) |>
  ggplot(aes(x = affpol_weighted, y = k_outgroup_ingroup)) +
  geom_point(alpha = 0.2, size = 0.8, shape = 21, fill = "grey40") +
  stat_smooth(
    method = "lm", formula = y ~ x,
    colour = "black", fill = "grey60",
    alpha = 0.15, linewidth = 0.85, se = TRUE
  ) +
  annotate(
    "text",
    x = -1, y = Inf,
    hjust = -0.05, vjust = 2.0,
    label = stats_label,
    size = 3.5, family = "Public Sans", colour = "grey20"
  ) +

  # Arrow along x-axis
  annotate(
    "segment",
    x = -1, xend = 2,
    y = -3.5, yend = -3.5,
    arrow = arrow(length = unit(0.15, "cm"), type = "closed"),
    colour = "grey20", linewidth = 0.4
  ) +
  # Label centered on the arrow
  annotate(
    "text",
    x = 0.7, y = -3.3,
    hjust = 0, vjust = 0,
    label = "More polarized",
    size = 3, family = "Public Sans", colour = "grey20"
  ) +
  
  coord_cartesian(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.18))) +
  labs(
    x     = "Affective Polarization",
    y     = "In-Group Favoritism"
  ) +
  theme(
    legend.position = "none",
    plot.margin = margin(t = 5, r = 10, b = 25, l = 5)  # extra bottom room
  )

p_scatter
```

```{r}
#| fig-width: 7
#| fig-height: 4
#| fig-dpi: 600
showtext_opts(dpi = 600)

fig_affpol <- p_fav_partisan + p_scatter +
  plot_annotation(
    title = "Study 2",
    theme = theme(plot.title = element_text(size = 18, face = "bold",
                                            family = "Public Sans"))
  ) &
  theme(legend.title = element_blank())


fig_affpol 

ggsave(here(fig_dir, "fig_affpol.png"),
       width = 7, height = 4, dpi = 600)
```

## Treatment Figures

```{r}
p_misperception_data <- survey_data |>
  count(condition, deviation) |>
  group_by(condition) |>
  mutate(prop = n / sum(n) * 100) |>
  ungroup()

condition_colors <- c("Control" = "grey90", "Treatment" = "#CC3311")

# Mean deviation (-1.57) mapped to discrete x position:
# levels are -3,-2,-1,0,1 → positions 1,2,3,4,5
# so -1.57 → -1.57 - (-3) + 1 = 2.43
mean_xpos <- 2.43

p_misperception <- ggplot(p_misperception_data, aes(x = factor(deviation), y = prop, fill = condition)) +
  # Highlight correct answer column
  annotate("rect",
           xmin = 3.5, xmax = 4.5, ymin = 0, ymax = Inf,
           fill = "forestgreen", alpha = 0.07) +
  geom_col(
    position = position_dodge(width = 0.75),
    width = 0.65, alpha = 0.6, color = "black"
  ) +
  # Add % over bar
  geom_text(
    aes(label = paste0(round(prop, 1), "%")),
    position = position_dodge(width = 0.75),
    vjust = -0.4,
    size = 3,
    family = "Public Sans",
    color = "grey20") +
  # Mean dashed vertical line
  geom_vline(
    xintercept = mean_xpos,
    linetype = "dashed", color = "grey30", linewidth = 0.6
  ) +
  # Correct answer label
  annotate("text",
           x = 4, y = 50,
           label = "Correct\nanswer", size = 3.5, color = "grey20",
           lineheight = 0.9, 
           family = "Public Sans") +
  scale_x_discrete(
    labels = c(
      "-3" = "Almost\nNever",
      "-2" = "Rarely",
      "-1" = "Sometimes",
       "0" = "Often",
       "1" = "Almost\nAlways"
    )
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.08)),
    breaks = seq(0, 50, 10),
    limits = c(0, 50),
    labels = function(x) paste0(x, "%")
  ) +
  scale_fill_manual(values = condition_colors) +
  labs(
    title = "(A) Perceptions of Out-Party Efforts",
    x = NULL,
    y = "Participants (%)",
    fill = NULL
  ) 

p_misperception
```

```{r}
treat_data_raw <- survey_data |>
  filter(deviation < 0) |>
  select(condition, k_ingroup, k_outgroup) |>
  standardise(select = c("k_ingroup", "k_outgroup"), center = 0) |> 
  mutate(across(c(k_ingroup, k_outgroup), as.numeric)) |>
  pivot_longer(
    cols      = c(k_ingroup, k_outgroup),
    names_to  = "beneficiary",
    values_to = "k"
  ) |>
  mutate(
    beneficiary = factor(
      recode(beneficiary, k_ingroup = "In-Group", k_outgroup = "Out-Group"),
      levels = c("In-Group", "Out-Group")
    ),
    condition = factor(condition, levels = c("Control", "Treatment"))
  )

treat_sum_raw <- treat_data_raw |>
  summarise(
    mean_k = mean(k, na.rm = TRUE),
    se_k   = sd(k, na.rm = TRUE) / sqrt(sum(!is.na(k))),
    .by    = c(condition, beneficiary)
  )

treat_signif_df <- tibble(
  group1      = c(0.825, 1.825),
  group2      = c(1.175, 2.175),
  label       = c("ns (b = \u22120.08)", "*** (b = \u22120.27)"),
  y_position  = c(
    max(treat_data_raw$k[treat_data_raw$beneficiary == "In-Group"],  na.rm = TRUE) * 1.05,
    max(treat_data_raw$k[treat_data_raw$beneficiary == "Out-Group"], na.rm = TRUE) * 1.05
  )
)
```

```{r}
#| fig-width: 5
#| fig-height: 3
#| fig-dpi: 600
p_treat <- ggplot() +
  geom_col(
    data     = treat_sum_raw,
    aes(x = beneficiary, y = mean_k, fill = condition),
    position = position_dodge(width = 0.7),
    color = "black", alpha = 0.6, width = 0.6
  ) +
  geom_jitter(
    data     = treat_data_raw,
    aes(x = beneficiary, y = k, fill = condition),
    shape    = 21, color = "black",
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.7),
    alpha = 0.3, size = 0.9,
    show.legend = FALSE
  ) +
  geom_errorbar(
    data     = treat_sum_raw,
    aes(x = beneficiary, ymin = mean_k - se_k, ymax = mean_k + se_k, group = condition),
    position = position_dodge(width = 0.7),
    width = 0.15, linewidth = 0.5,
    show.legend = FALSE
  ) +
  geom_signif(
    data        = treat_signif_df,
    aes(xmin = group1, xmax = group2, annotations = label, y_position = y_position),
    manual      = TRUE,
    tip_length  = 0,
    size        = 0.8,
    textsize    = 3.5,
    family      = "Public Sans",
    color       = "black"
  ) +
  scale_fill_manual(values = condition_colors, name = NULL) +
  scale_y_continuous(
    limits = c(0, NA),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(title = "(B) Correcting Misperceptions", x = NULL, y = "Effort Discounting (K)") 

p_treat
```

```{r}
#| fig-width: 10
#| fig-height: 4.5
#| fig-dpi: 600
showtext_opts(dpi = 600)

fig_treat <- wrap_plots(
    p_misperception, p_treat, guide_area(),
    design  = c(area(1,1,1,3), area(1,4,1,5), area(2,1,2,5)),
    heights = c(10, 1),
    guides  = "collect"
  )  &
  theme(legend.title = element_blank(), legend.position = "bottom")

fig_treat

ggsave(here(fig_dir, "fig_treat.png"),
       width = 10, height = 4.5, dpi = 600)
```




## Experiment Descriptives

### 1. Random Assignment by Party

```{r}
survey_data |> 
  select(Condition = condition,
         Party = party) |> 
  droplevels() |> 
  datasummary_crosstab(Condition ~ Party, 
                       data = _)
```

### 2. Misperceptions?

```{r}
# Build summary table
underest_table <- survey_data |>
  summarise(
    N_total = n(),
    N_under = sum(deviation < 0),
    Pct_under = mean(deviation < 0) * 100,
    .by = condition
  ) |>
  bind_rows(
    survey_data |>
      summarise(
        condition = factor("Overall"),
        N_total = n(),
        N_under = sum(deviation < 0),
        Pct_under = mean(deviation < 0) * 100
      )
  ) |>
  arrange(condition)

# Render with tinytable
tt(underest_table,
   digits = 1,
   caption = "Participants underestimating out-party willingness (deviation < 0)") |>
  setNames(c("Condition", "N (Total)", "N (Underestimated)", "% Underestimated")) |>
  format_tt(j = 4, sprintf = "%.1f%%")
```


#### Democrats

```{r}
survey_data |> 
  filter(party_preference == "Democrat",
         condition == "Treatment") |> 
  select(treat_est_dem_01) |> 
  datasummary_skim()
```

#### Republicans

```{r}
survey_data |> 
  filter(party_preference == "Republican",
         condition == "Treatment") |> 
  select(treat_est_rep_01) |> 
  datasummary_skim()
```

### 3. Manipulation Checks

```{r}
survey_data |> 
  filter(party == "Democrat",
         condition == "Treatment") |> 
  select(treat_dem_surprise, treat_dem_credible) |> 
  mutate(across(c(everything()), as.numeric)) |> 
  datasummary_skim()
```

```{r}
survey_data |> 
  filter(party == "Republican",
         condition == "Treatment") |> 
  select(treat_rep_surprise, treat_rep_credible) |> 
  mutate(across(c(everything()), as.numeric)) |> 
  datasummary_skim()
```

### Computational Model Fit

Read relevant files
```{r}
fit_data_dir <- here("01_data", "models", "matlab_fit")
fit <- read.csv(here(fit_data_dir, "PM_model_fit_statistics_full_democraticeffortdata.csv"))
params <- read.csv(here(fit_data_dir, "EM_fit_parameters_full_democraticeffortdata_ms_two_k_one_beta_hyperbolic_labeled.csv"))
```

```{r}
# Font for greek symbols
font_add_google("Noto Sans", "noto")
```

Prepare data
```{r}
fit_plot_dat <- fit |>
  mutate(
    delta_bic = bicint - min(bicint),
    model_label = case_when(
      model_name == "ms_two_k_one_beta" ~ "2K1β\nParabolic",
      model_name == "ms_two_k_one_beta_linear" ~ "2K1β\nLinear",
      model_name == "ms_two_k_one_beta_hyperbolic" ~ "2K1β\nHyperbolic",
      model_name == "ms_two_k_two_beta" ~ "2K2β\nParabolic",
      model_name == "ms_two_k_two_beta_linear" ~ "2K2β\nLinear",
      model_name == "ms_two_k_two_beta_hyperbolic" ~ "2K2β\nHyperbolic",
      TRUE ~ model_name
    )
  ) |>
  arrange(delta_bic) |>
  mutate(model_label = factor(model_label, levels = model_label))

winner_row <- fit |>
  filter(model_name == "ms_two_k_one_beta_hyperbolic")

winner_text <- paste0(
  "Winning model\n\n",
  "2K1β hyperbolic\n\n",
  "BICint = ", format(round(winner_row$bicint, 0), big.mark = ","), "\n",
  "ΔBIC = ", round(winner_row$relbic, 0), "\n",
  "XP = ", sprintf("%.2f", winner_row$xp)
)
```

```{r}
#| fig-width: 4.5
#| fig-height: 3.5
#| fig-dpi: 600
showtext_opts(dpi = 600)

fit_plot <- ggplot(fit_plot_dat, aes(x = model_label, y = delta_bic, fill = model_label)) +
  geom_col(width = 0.68, colour = "black") +
  geom_text(
    aes(label = round(delta_bic, 1)),
    vjust = -0.35,
    size = 3.5
  ) +
  annotate(
    "label",
    x    = 0.55,
    y    = max(fit_plot_dat$delta_bic) * 1.12,
    label = winner_text,
    hjust = 0,
    vjust = 1,
    size        = 3,
    linewidth = 0.35,
    fill        = "grey96",
    colour      = "grey25",
    fontface    = "bold",
    family      = "noto",
    label.r     = unit(0.2, "lines"),
    label.padding = unit(0.45, "lines")
  ) +
  coord_cartesian(
    ylim = c(0, max(fit_plot_dat$delta_bic) * 1.12),
    clip = "off"
  ) +
  scale_fill_manual(
    values = rev(RColorBrewer::brewer.pal(9, "OrRd")[3:9]),
    guide  = "none") +
  labs(
    title = NULL,
    x = NULL,
    y = expression(Delta * "BIC"[int])
  ) +
  theme_nice(base_size = 11) +
  theme(
    axis.title.y = element_text(family = "noto"),
    axis.text.x  = element_text(
      family     = "noto",
      angle      = 30,
      hjust      = 1,
      vjust      = 1,
      lineheight = 0.9
    )
  )

fit_plot
```

```{r}
reward_level <- 4

curve_df <- combined_data |>
  summarise(
    `In-group`  = median(k_ingroup,  na.rm = TRUE),
    `Out-group` = median(k_outgroup, na.rm = TRUE)
  ) |>
  pivot_longer(everything(), names_to = "beneficiary", values_to = "K") |>
  crossing(effort = seq(0, 3, length.out = 200)) |>
  mutate(
    subjective_value = reward_level / (1 + K * effort),
    beneficiary      = factor(beneficiary, levels = c("In-group", "Out-group"))
  )
```

```{r}
#| fig-width: 4.5
#| fig-height: 3.5
#| fig-dpi: 600
showtext_opts(dpi = 600)

curve_plot <- ggplot(curve_df, aes(x = effort, y = subjective_value, colour = beneficiary)) +
  geom_line(linewidth = 1.2) +
  scale_colour_manual(
    values = c("In-group" = "#1B9E91", "Out-group" = "#E97B32")
  ) +
  scale_x_continuous(limits = c(0, 3), breaks = 0:3) +
  labs(
    x      = "Effort level",
    y      = "Subjective value",
    colour = NULL
  ) +
  guides(
    colour = guide_legend(override.aes = list(linewidth = 2))
  ) +
  theme_nice(base_size = 13) +
  theme(legend.position = "bottom")

curve_plot
```

```{r}
#| fig-width: 4.5
#| fig-height: 7
#| fig-dpi: 600
showtext_opts(dpi = 600)

(fit_plot + theme(plot.margin = margin(t = 5, r = 5, b = 20, l = 5))) / curve_plot +
  plot_annotation(
    title = "(B)",
    theme = theme(plot.title = element_text(face = "bold", size = 12))
  ) 

ggsave(
  here(fig_dir, "fit_curve.svg"),
  height = 7,
  width  = 4.5,
  dpi    = 600
)
```






















