---
title: "Democratic Effort Study 2 Models"
format:
  html:
    theme: paper
    code-tools: true
    code-fold: true
    toc: true
    toc-fold: false
    toc-location: right
    toc-depth: 3
    toc-title: "Contents"
embed-resources: true
warning: false
---

```{r}
#| include: false
# set error messages to English
Sys.setenv(LANG = "en")

knitr::opts_chunk$set(
  fig.align = "center", 
  fig.retina = 3,
  fig.width = 6, 
  fig.height = (6 * 0.618),
  out.width = "90%", 
  collapse = FALSE
)

options(
  digits = 4,
  width = 300,
  dplyr.summarise.inform = FALSE)
```

```{r}
#| include: false
#| warning: true
#| message: true
if (!require("groundhog", quietly = TRUE)) {
  install.packages("groundhog")
}
library("groundhog")

# use groundhog to install and load packages
pkgs <- c("here",         # System path management
          "datawizard",   # Data preparations (e.g., standardise)
          "tidyverse",    # ggplot, dplyr, %>%, and friends
          "modelsummary", # Data and model summaries with tables and plots
          "tinytable",    # Lightweight package to create tables
          "afex",         # Interface for (general) linear mixed models
          "estimatr",     # Fast estimators with robust standard errors
          "BayesFactor",  # R package for BayesFactors
          "mgcv",         # Generalized additive models
          "marginaleffects", # Compute marginal means, slopes, predictions, etc.
          "ggdist",       # Additional plot elements
          "patchwork",    # Combine multiple ggplots
          "showtext"      # Additional fonts for plots etc
          )

groundhog.library(pkgs, "2026-01-01")
```

```{r}
#| include: false
showtext_auto()
showtext_opts(dpi = 600)

font_add_google("Public Sans", "Public Sans")
font_add_google("Public Sans", "Public Sans Light", regular.wt = 300)
font_add_google("Public Sans", "Public Sans Semibold", regular.wt = 600)

# Custom ggplot theme
theme_nice <- function() {
  # General text family and base size
  theme_minimal(base_family = "Public Sans", base_size = 14) +
    theme(
      # Remove minor grid lines for cleaner look
      panel.grid.minor = element_blank(),
      # Customize major grid lines color
      panel.grid.major = element_blank(),
      # Plot title: bold and custom relative size
      plot.title = element_text(family = "Public Sans", face = "bold", size = rel(1.1), margin = margin(b = 10)),
      # Subtitle and caption: lighter weight 
      plot.subtitle = element_text(family = "Public Sans Light", face = "plain"),
      plot.caption = element_text(family = "Public Sans Light", face = "plain"),
      # Axis text: Relative size
      axis.text = element_text(size = rel(0.9)),
      # Axis titles: Semibold and custom relative size
      axis.title = element_text(family = "Public Sans Semibold", size = rel(0.9)),
      axis.title.x = element_text(hjust = 0.5),
      axis.title.y = element_text(hjust = 0.5),
      # Axis lines
      axis.line = element_line(color = "grey25", linewidth = 0.8),
      # Facet labels: bold text on grey background
      strip.text = element_text(
        family = "Public Sans Semibold", face = "plain",
        size = rel(1),
        hjust = 0.5
      ),
      strip.background = element_rect(fill = "grey95", color = "grey25", linewidth = 0.8), 
      # Legend styling: compact and top-positioned
      legend.title = element_text(family = "Public Sans Semibold", size = rel(0.8)),
      legend.text = element_text(size = rel(0.8)),
      legend.position = "top",
      legend.justification = "center",
      legend.title.position = "left",
      legend.margin = margin(l = 0, t = 0)
    )
}

theme_set(theme_nice())
update_geom_defaults("text", list(family = "Public Sans"))
update_geom_defaults("label", list(family = "Public Sans"))
```

```{r}
#| include: false
data_dir <- here("01_data", "final") 
table_dir <- here("02_output", "tables", "prereg") 
fig_dir <- here("02_output", "figures", "prereg") 

dir.create(table_dir, recursive = TRUE, showWarnings = TRUE)
dir.create(fig_dir, recursive = TRUE, showWarnings = TRUE)
```

## Load and prepare data

```{r}
load(here(data_dir, "demeffort_survey_data.RData")) 
load(here(data_dir, "demeffort_combined_data.RData"))
```

```{r}
model_data <- survey_data |>
  standardize(
    select = c(
      "k_rural", "k_urban", "k_ingroup", "k_outgroup", "k_mean",
      "k_outgroup_ingroup", "beta", "party_importance", "affpol_raw",
      "affpol_weighted", "valence_democrats", "valence_republicans",
      "arousal_democrats", "arousal_republicans", "democratic_support", "freefair"
    )
  ) 

model_data_dev <- survey_data |> 
  filter(deviation < 0) |> 
  standardize(
    select = c(
      "k_rural", "k_urban", "k_ingroup", "k_outgroup", "k_mean",
      "k_outgroup_ingroup", "beta", "party_importance", "affpol_raw",
      "affpol_weighted", "valence_democrats", "valence_republicans",
      "arousal_democrats", "arousal_republicans", "democratic_support", "freefair"
    )
  ) 
```

# Democratic Effort Task Study 2 Pre-Registered Models

## H1: Deviation 

H1. Participants' estimates of how willing out-party supporters are to exert effort to support democratic engagement in areas that have traditionally leaned toward their own party (e.g., Democrats estimating Republican willingness to work for urban areas) will, on average, be lower than the actual level of effort willingness presented as corrective feedback in the treatment condition.

```{r}
h1a <- lm_robust(deviation ~ 1, se_type = "HC2", data = survey_data)
summary(h1a)
```

```{r}
h1b <- lm_robust(deviation ~ condition, se_type = "HC2", data = survey_data)
summary(h1b)
```

```{r}
h1.table <- modelsummary(
  list("H1a: Average Deviation" = h1a,
       "H1b: + Condition"    = h1b),
  shape      = model + term ~ statistic,
  title      = "Table H1: Baseline Misperceptions (Deviation from Correct Answer)",
  coef_rename = c(
    "(Intercept)"        = "",
    "conditionTreatment" = "Treatment vs. Control"
  ),
  estimate   = "{estimate}{stars}",
  statistic  = c("{std.error}", "conf.int", "p.value"),
  conf_level = .95,
  stars      = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  fmt        = 3,
  notes      = list("p-values: * < .05, ** < .01, *** < .001"),
  gof_omit   = ".*"
)

h1.table
h1.table |> save_tt(here(table_dir, "h1.table.docx"), overwrite = TRUE)
```

## H2A: Treatment Effects on In Group Favouritism

H2A. Among participants who underestimate out-party effort willingness (i.e., whose estimate falls below the actual value presented as corrective feedback), those in the treatment condition (vs. control) will show a smaller difference in their motivation to exert effort for the in-group versus the out-group.

```{r}
h2a <- lm_robust(
  k_outgroup_ingroup ~ condition,
  se_type = "HC2",
  data = model_data_dev)

summary(h2a)
```

```{r}
h2a.table <- modelsummary(
  list("H2A" = h2a),
  shape = term ~ model + statistic,
  title      = "Table H2A: ATE on In-Group Favouritism",
  coef_rename = c(
    "(Intercept)"        = "Intercept (Control)",
    "conditionTreatment" = "Treatment vs. Control"
  ),
  estimate   = "{estimate}{stars}",
  statistic  = c("{std.error}", "conf.int", "p.value"),
  conf_level = .95,
  stars      = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  fmt        = 3,
  notes      = list("p-values: * < .05, ** < .01, *** < .001"),
  gof_omit   = ".*"
)

h2a.table
h2a.table |> save_tt(here(table_dir, "h2a.table.docx"), overwrite = TRUE)
```

## H2B: Treatment Effect specific to the Outgroup?

H2B. (Follow-up to H2A) This reduction in the treatment vs. control will be driven by a decrease in effort discounting (increase in motivation) for the out-group rather than a change in effort discounting/motivation for the in-group (among participants with misperceptions).

Reshape the data for the linear mixed model:
```{r}
model_data_dev_long <- model_data_dev |>
  select(PID, condition, deviation, k_ingroup, k_outgroup) |>
  pivot_longer(
    cols      = c(k_ingroup, k_outgroup),
    names_to  = "benefit",
    values_to = "k"
  ) |>
  mutate(
    benefit = factor(benefit,
                     levels = c("k_ingroup", "k_outgroup"),
                     labels = c("In-group", "Out-group"))
  )
```

Fit the linear-mixed model:
```{r}
h2b <- mixed(
  k ~ benefit * condition + (1 | PID),
  data = model_data_dev_long
)

h2b
summary(h2b)
```

::: {.callout-note}
## Linear Mixed Model was misspecified in pre-registration

Not a problem but note the deviation. We can't use a random slope as benefit only has two levels:
Error: number of observations (=1100) <= number of random effects (=1100) for term (benefit | PID); the random-effects parameters and the residual variance (or scale parameter) are probably unidentifiable
:::

```{r}
# Extract merMod object
h2b.mer <- h2b$full_model

# Pairwise condition comparisons by benefit level
h2b.com <- avg_comparisons(h2b.mer,
                                   variables = list(condition = "pairwise"),
                                   by = "benefit",
                                   re.form = NA) |>
  as_tibble() |>
  select(Contrast = contrast,
         Benefit = benefit,
         Estimate = estimate,
         SE = std.error,
         "t" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high) |>
  mutate(
    across(where(is.numeric) & !matches("p"), ~ round(.x, 3)),
    p = ifelse(round(p, 3) == 0, "< .001", as.character(round(p, 3)))
  )

h2b.com |> tt()
```

```{r}
h2b.hyp <- avg_comparisons(h2b.mer,
                            variables = "condition",
                            by = "benefit",
                            re.form = NA,
                            hypothesis = ~pairwise) |>
  as_tibble() |>
  select(Hypothesis = hypothesis,
         Estimate = estimate,
         SE = std.error,
         "t" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high) |>
  mutate(
    across(where(is.numeric) & !matches("p"), ~ round(.x, 3)),
    p = ifelse(round(p, 3) == 0, "< .001", as.character(round(p, 3)))
  )

h2b.hyp |> tt(caption = "H2B: Test of whether the treatment effect on effort discounting 
     differs between in-group and out-group beneficiaries")
```

## H3A: In-Group Favouritism ~ Affective Polarisation 

H3A. Affective polarization will be positively associated with greater differences in the motivation to exert effort for the in-group versus the out-group.

```{r}
h3a <- lm_robust(
  k_outgroup_ingroup ~ affpol_weighted + condition,
  se_type = "HC2",
  data    = model_data
)

summary(h3a)
```

```{r}
h3a.table <- modelsummary(
  list("H3A" = h3a),
  shape      = term ~ model + statistic,
  title      = "Table H3A: Affective Polarization and In-Group Favouritism",
  coef_rename = c(
    "(Intercept)"        = "Intercept (Control)",
    "affpol_weighted"    = "Affective Polarization",
    "conditionTreatment" = "Treatment vs. Control"
  ),
  estimate   = "{estimate}{stars}",
  statistic  = c("{std.error}", "conf.int", "p.value"),
  conf_level = .95,
  stars      = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  fmt        = 3,
  notes      = list("p-values: * < .05, ** < .01, *** < .001"),
  gof_omit   = ".*"
)

h3a.table
h3a.table |> save_tt(here(table_dir, "h3a.table.docx"), overwrite = TRUE)
```

## H3B: Deviation ~ Affective Polarisation

H3B. Affective polarization will be positively associated with stronger baseline misperceptions about how willing out-party supporters are to exert effort to support democratic engagement in areas that have traditionally leaned toward their own party.

```{r}
h3b <- lm_robust(
  deviation ~ affpol_weighted + condition,
  se_type = "HC2",
  data    = model_data
)

summary(h3b)
```

```{r}
h3b.table <- modelsummary(
  list("H3B" = h3b),
  shape      = term ~ model + statistic,
  title      = "Table H3B: Affective Polarization and Baseline Misperceptions",
  coef_rename = c(
    "(Intercept)"        = "Intercept (Control)",
    "affpol_weighted"    = "Affective Polarization",
    "conditionTreatment" = "Treatment vs. Control"
  ),
  estimate   = "{estimate}{stars}",
  statistic  = c("{std.error}", "conf.int", "p.value"),
  conf_level = .95,
  stars      = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  fmt        = 3,
  notes      = list("p-values: * < .05, ** < .01, *** < .001"),
  gof_omit   = ".*"
)

h3b.table
h3b.table |> save_tt(here(table_dir, "h3b.table.docx"), overwrite = TRUE)
```

## Exploratory: In-Group Favouritism ~ Deviation

```{r}
h3c <- lm_robust(
  k_outgroup_ingroup ~ deviation + condition,
  se_type = "HC2",
  data    = model_data_dev
)

summary(h3c)
```

```{r}
h3c.table <- modelsummary(
  list("H3C" = h3c),
  shape      = term ~ model + statistic,
  title      = "Table H3C (Exploratory): Baseline Misperceptions and In-Group Favouritism",
  coef_rename = c(
    "(Intercept)"        = "Intercept (Control)",
    "deviation"          = "Deviation (Misperception Strength)",
    "conditionTreatment" = "Treatment vs. Control"
  ),
  estimate   = "{estimate}{stars}",
  statistic  = c("{std.error}", "conf.int", "p.value"),
  conf_level = .95,
  stars      = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  fmt        = 3,
  notes      = list("p-values: * < .05, ** < .01, *** < .001"),
  gof_omit   = ".*"
)

h3c.table
h3c.table |> save_tt(here(table_dir, "h3c.table.docx"), overwrite = TRUE)
```

## H4: Conditional Treatment Effects (AffPol)

We have two competing hypotheses on how affective polarisation might affect the effectiveness of the corrective feedback presented in the treatment group. 
H4A. The corrective feedback will be less effective at reducing in-group versus out-group effort differences among participants higher in affective polarization (for example, because more polarized participants may be more resistant to updating their views of the out-party or may find the feedback less credible).
H4B. The corrective feedback will be more effective at reducing in-group versus out-group effort differences among participants higher in affective polarization (for example, because more polarized participants may hold stronger misperceptions that leave more room for correction).

```{r}
h4 <- lm_robust(
  k_outgroup_ingroup ~ condition * affpol_weighted,
  se_type = "HC2",
  data    = model_data_dev
)

summary(h4)
```

```{r}
h4.table <- modelsummary(
  list("H4" = h4),
  shape      = term ~ model + statistic,
  title      = "Table H4: Moderation of Treatment Effect by Affective Polarization",
  coef_rename = c(
    "(Intercept)"                        = "Intercept (Control)",
    "conditionTreatment"                 = "Treatment vs. Control",
    "affpol_weighted"                    = "Affective Polarization",
    "conditionTreatment:affpol_weighted" = "Treatment × Affective Polarization"
  ),
  estimate   = "{estimate}{stars}",
  statistic  = c("{std.error}", "conf.int", "p.value"),
  conf_level = .95,
  stars      = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  fmt        = 3,
  notes      = list("p-values: * < .05, ** < .01, *** < .001"),
  gof_omit   = ".*"
)

h4.table
h4.table |> save_tt(here(table_dir, "h4.table.docx"), overwrite = TRUE)
```










