---
title: "Democratic Effort Study 2: Follow-Up Regressions"
format:
  html:
    theme: paper
    code-tools: true
    code-fold: true
    toc: true
    toc-fold: false
    toc-location: right
    toc-depth: 3
    toc-title: "Contents"
embed-resources: true
warning: false
---

```{r}
#| include: false
# set error messages to English
Sys.setenv(LANG = "en")

knitr::opts_chunk$set(
  fig.align = "center", 
  fig.retina = 3,
  fig.width = 6, 
  fig.height = (6 * 0.618),
  out.width = "90%", 
  collapse = FALSE
)

options(
  digits = 4,
  width = 300,
  dplyr.summarise.inform = FALSE)
```

```{r}
#| include: false
#| warning: true
#| message: true
if (!require("groundhog", quietly = TRUE)) {
  install.packages("groundhog")
}
library("groundhog")

# use groundhog to install and load packages
pkgs <- c("here",         # System path management
          "datawizard",   # Data preparations (e.g., standardise)
          "tidyverse",    # ggplot, dplyr, %>%, and friends
          "modelsummary", # Data and model summaries with tables and plots
          "tinytable",    # Lightweight package to create tables
          "afex",         # Interface for (general) linear mixed models
          "estimatr",     # Fast estimators with robust standard errors
          "broom.mixed",  # Get tidy model objects
          "BayesFactor",  # R package for BayesFactors
          "mgcv",         # Generalized additive models
          "marginaleffects", # Compute marginal means, slopes, predictions, etc.
          "ggdist",       # Additional plot elements
          "patchwork",    # Combine multiple ggplots
          "showtext"      # Additional fonts for plots etc
          )

groundhog.library(pkgs, "2026-01-01")
```

```{r}
#| include: false
showtext_auto()
showtext_opts(dpi = 600)

font_add_google("Public Sans", "Public Sans")
font_add_google("Public Sans", "Public Sans Light", regular.wt = 300)
font_add_google("Public Sans", "Public Sans Semibold", regular.wt = 600)

# Custom ggplot theme
theme_nice <- function() {
  # General text family and base size
  theme_minimal(base_family = "Public Sans", base_size = 14) +
    theme(
      # Remove minor grid lines for cleaner look
      panel.grid.minor = element_blank(),
      # Customize major grid lines color
      panel.grid.major = element_blank(),
      # Plot title: bold and custom relative size
      plot.title = element_text(family = "Public Sans", face = "bold", size = rel(1.1), margin = margin(b = 10)),
      # Subtitle and caption: lighter weight 
      plot.subtitle = element_text(family = "Public Sans Light", face = "plain"),
      plot.caption = element_text(family = "Public Sans Light", face = "plain"),
      # Axis text: Relative size
      axis.text = element_text(size = rel(0.9)),
      # Axis titles: Semibold and custom relative size
      axis.title = element_text(family = "Public Sans Semibold", size = rel(0.9)),
      axis.title.x = element_text(hjust = 0.5),
      axis.title.y = element_text(hjust = 0.5),
      # Axis lines
      axis.line = element_line(color = "grey25", linewidth = 0.8),
      # Facet labels: bold text on grey background
      strip.text = element_text(
        family = "Public Sans Semibold", face = "plain",
        size = rel(1),
        hjust = 0.5
      ),
      strip.background = element_rect(fill = "grey95", color = "grey25", linewidth = 0.8), 
      # Legend styling: compact and top-positioned
      legend.title = element_text(family = "Public Sans Semibold", size = rel(0.8)),
      legend.text = element_text(size = rel(0.8)),
      legend.position = "top",
      legend.justification = "center",
      legend.title.position = "left",
      legend.margin = margin(l = 0, t = 0)
    )
}

theme_set(theme_nice())
update_geom_defaults("text", list(family = "Public Sans"))
update_geom_defaults("label", list(family = "Public Sans"))
```

```{r}
#| include: false
data_dir <- here("01_data", "final") 
table_dir <- here("02_output", "tables", "follow_up") 
fig_dir <- here("02_output", "figures", "follow_up") 
model_dir <- here("02_output", "models", "follow_up") 

dir.create(table_dir, recursive = TRUE, showWarnings = TRUE)
dir.create(fig_dir, recursive = TRUE, showWarnings = TRUE)
dir.create(model_dir, recursive = TRUE, showWarnings = TRUE)
```

# Democratic Effort Task Regressions

## Load and prepare data

```{r}
load(here(data_dir, "demeffort_survey_data.RData")) 
load(here(data_dir, "demeffort_combined_data.RData"))
```

```{r}
participant_vars <- c(
  "k_rural", "k_urban", "k_ingroup", "k_outgroup", "k_mean",
  "k_outgroup_ingroup", "beta", "party_importance", "affpol_raw",
  "affpol_weighted", "valence_democrats", "valence_republicans",
  "arousal_democrats", "arousal_republicans", "democratic_support", "freefair"
)

# Wide datasets — standardized
model_data_wide <- survey_data |>
  standardize(select = participant_vars)

model_data_wide_dev <- survey_data |>
  filter(deviation < 0) |>
  standardize(select = participant_vars)

# Long datasets, drop raw participant vars, merge in standardized ones
model_data_long <- combined_data |>
  select(-any_of(participant_vars)) |>
  inner_join(
    model_data_wide |> select(PID, all_of(participant_vars)),
    by = "PID"
  ) |>
  mutate(rew = factor(rew, levels = c(4, 12, 20))) |>
  filter(!is.na(decision)) |>
  droplevels()

model_data_long_dev <- combined_data |>
  select(-any_of(participant_vars)) |>
  inner_join(
    model_data_wide_dev |> select(PID, all_of(participant_vars)),
    by = "PID"
  ) |>
  mutate(rew = factor(rew, levels = c(4, 12, 20))) |>
  filter(!is.na(decision)) |>
  droplevels()
```

## Generalised linear mixed models

Modelling strategy:

1.  Fix maximum models (Barr et al., 2013)

2.  If models don't converge:

    1.  Remove the correlations between random slope and intercept

    2.  Remove interactions between random slope effects

    3.  Remove random slope effects with a variance close to 0

    4.  Run models only including a random intercept

### Model 1: Decision ~ Reward x Effort x Condition x Benefit

```{r}
# number of observations (=8933) < number of random effects (=9900)
# m1.full <- mixed(decision ~ condition * benefit * eff * rew + (1 + benefit * eff * rew | PID),
#             data = model_data_long_dev,
#             family = binomial,
#             method = "LRT",
#             control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e5)),
#             return = "mixed",
#             expand_re = T)
# 
# m1.full
# summary(m1.full)
```

Remove correlations between random terms
```{r}
# boundary (singular) fit: see help('isSingular')
# m1.nocorr <- mixed(decision ~ condition * benefit * eff * rew + (1 + benefit * eff * rew || PID),
#             data = model_data_long_dev,
#             family = binomial,
#             method = "LRT",
#             control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e5)),
#             return = "merMod",
#             expand_re = T)
# 
# m1.nocorr
# summary(m1.nocorr)
```

Remove interactions between random terms
```{r}
# boundary (singular) fit: see help('isSingular')
# m1.noint <- mixed(decision ~ condition * benefit * eff * rew + (1 + benefit + eff + rew || PID),
#             data = model_data_long_dev,
#             family = binomial,
#             method = "LRT",
#             control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e5)),
#             return = "merMod",
#             expand_re = T)
# 
# m1.noint 
# summary(m1.noint)
```

Remove random slopes close to 0 (re-introduce correlation between random slope and random intercept)
```{r}
m1_path <- here(model_dir, "m1_glmm_decision.rds")

if (!file.exists(m1_path)) {
  m1 <- mixed(
    decision ~ condition * benefit * eff * rew + 
               (1 + benefit | PID),
    data = model_data_long_dev,
    family = binomial,
    method = "LRT",
    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)),
    return = "mixed",
    expand_re = TRUE
  )
  saveRDS(m1, m1_path)
} else {
  m1 <- readRDS(m1_path)
}
```

```{r}
anova_table <- m1$anova_table |>
  as_tibble(rownames = "Effect") |>
  select(
    Effect,
    df   = `Chi Df`,
    Chi2 = Chisq,
    p    = `Pr(>Chisq)`
  ) |>
  mutate(
    Effect = Effect |>
      str_replace_all("benefit",   "Beneficiary") |>
      str_replace_all("rew",       "Reward") |>
      str_replace_all("eff",       "Effort") |>
      str_replace_all("condition", "Condition") |>
      str_replace_all(":", " × "),
    Chi2 = sprintf("%.2f", Chi2),
    p    = ifelse(p < .001, "< .001", sprintf("%.3f", p))
  )

anova_table |>
  tt() |>
  save_tt(here(table_dir, "anova_table.docx"), overwrite = TRUE)

anova_table |>
  tt()
```

```{r}
contrasts(m1$data$condition)
contrasts(m1$data$benefit)
contrasts(m1$data$eff)
contrasts(m1$data$rew)
```

```{r}
m1.mer <- m1$full_model

relabel_terms <- function(x) {
  x |>
    str_replace_all("condition1", "Control") |>
    str_replace_all("benefit1",   "In-group") |>
    str_replace_all("eff1",       "Effort Level 1") |>
    str_replace_all("eff2",       "Effort Level 2") |>
    str_replace_all("rew1",       "4 Credits") |>
    str_replace_all("rew2",       "12 Credits") |>
    str_replace_all(":",          " \u00d7 ")
}

fe <- tidy(m1.mer, effects = "fixed", conf.int = TRUE) |>
  mutate(
    term = relabel_terms(term),
    term = case_when(
      term == "(Intercept)"                           ~ term,
      str_detect(term, "\u00d7")                      ~ term,
      term == "Control"                               ~ "Condition: Control",
      term == "In-group"                              ~ "Beneficiary: In-group",
      term %in% c("Effort Level 1", "Effort Level 2") ~ paste0("Effort: ", term),
      term %in% c("4 Credits", "12 Credits")          ~ paste0("Reward: ", term),
      TRUE                                            ~ term
    )
  ) |>
  transmute(
    Term     = term,
    Estimate = sprintf("%.3f", estimate),
    SE       = sprintf("%.3f", std.error),
    z        = sprintf("%.2f", statistic),
    `95% CI` = paste0("[", sprintf("%.3f", conf.low), ", ", sprintf("%.3f", conf.high), "]"),
    p        = ifelse(p.value < .001, "< .001", sprintf("%.3f", p.value))
  )

re <- tidy(m1.mer, effects = "ran_pars") |>
  mutate(
    term = term |>
      str_replace_all("re1\\.benefit1",                  "In-group") |>
      str_replace_all("sd__\\(Intercept\\)",             "SD (Intercept)") |>
      str_replace_all("sd__In-group",                    "SD (In-group)") |>
      str_replace_all("cor__\\(Intercept\\)\\.In-group", "Cor (Intercept, In-group)")
  ) |>
  transmute(
    Term     = term,
    Estimate = sprintf("%.3f", estimate),
    SE       = "",
    z        = "",
    `95% CI` = "",
    p        = ""
  )

n_fe <- nrow(fe)

lmm_table <- bind_rows(fe, re) |>
  tt() |>
  group_tt(
    i = list(
      "Fixed Effects"                = 1,
      "Random Effects (Participant)" = n_fe + 1
    )
  ) |>
  style_tt(i = c(1, n_fe + 2), bold = TRUE)

lmm_table

lmm_table |>
  save_tt(here(table_dir, "lmm_table.docx"), overwrite = TRUE)
```

### Model-based predictions

```{r}
# 1. Predicted probabilities by beneficiary (replicates Study 1 finding)
m1.ben.means <- avg_predictions(m1.mer, 
                                variables = "benefit",
                                re.form = NA) |> 
  as_tibble() |> 
  select(Beneficiary = benefit, 
         Estimate = estimate, 
         SE = std.error,
         "z" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high)

# 2. Predicted probabilities by condition (the main treatment effect)
m1.cond.means <- avg_predictions(m1.mer, 
                                 variables = "condition",
                                 re.form = NA) |> 
  as_tibble() |> 
  select(Condition = condition, 
         Estimate = estimate, 
         SE = std.error,
         "z" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high)

# 3. Predicted probabilities by condition × beneficiary (the interaction)
m1.cond.ben.means <- avg_predictions(m1.mer, 
                                     variables = c("condition", "benefit"),
                                     re.form = NA) |> 
  as_tibble() |> 
  select(Condition = condition,
         Beneficiary = benefit, 
         Estimate = estimate, 
         SE = std.error,
         "z" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high)

m1.means <- bind_rows(
  m1.ben.means |> mutate(Panel = "Beneficiary"),
  m1.cond.means |> mutate(Panel = "Condition"),
  m1.cond.ben.means |> mutate(Panel = "Condition × Beneficiary")
) |> 
  mutate(
    across(where(is.numeric) & !matches("p"), ~ round(.x, 3)),
    p = ifelse(round(p, 3) == 0, "< .001", as.character(round(p, 3)))
  )

m1.means |> 
  tt() |> 
  save_tt(here(table_dir, "model_means.docx"), overwrite = TRUE)

m1.means |> tt()
```

```{r}
# Model-based comparisons
# 1. Beneficiary pairwise (replicates Study 1)
m1.ben.comparisons <- avg_comparisons(m1.mer, 
                                       variables = list(benefit = "pairwise"),
                                       re.form = NA) |> 
  as_tibble() |> 
  mutate(Panel = "Beneficiary")

# 2. Condition pairwise (overall treatment effect)
m1.cond.comparisons <- avg_comparisons(m1.mer, 
                                        variables = list(condition = "pairwise"),
                                        re.form = NA) |> 
  as_tibble() |> 
  mutate(Panel = "Condition")

# 3. Condition effect within each beneficiary (the interaction follow-up)
m1.cond.by.ben <- avg_comparisons(m1.mer, 
                                   variables = list(condition = "pairwise"),
                                   by = "benefit",
                                   re.form = NA) |> 
  as_tibble() |> 
  mutate(Panel = "Condition within Beneficiary")

m1.comparisons <- bind_rows(
  m1.ben.comparisons,
  m1.cond.comparisons,
  m1.cond.by.ben
) |> 
  select(Panel,
         Contrast = contrast,
         Beneficiary = any_of("benefit"),
         Estimate = estimate, 
         SE = std.error,
         "z" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high) |> 
  mutate(
    across(where(is.numeric) & !matches("p"), ~ round(.x, 3)),
    p = ifelse(round(p, 3) == 0, "< .001", as.character(round(p, 3)))
  )

m1.comparisons |> 
  tt() |> 
  save_tt(here(table_dir, "model_comparisons.docx"), overwrite = TRUE)

m1.comparisons |> tt()
```

## Comparing Over-, Accurate-, and Underestimators

```{r}
model_data_wide <- model_data_wide |>
  mutate(
    estimation_group = case_when(
      deviation < 0  ~ "Underestimators",
      deviation == 0 ~ "Accurate",
      deviation > 0  ~ "Overestimators"
    ) |>
      factor(levels = c("Underestimators", "Accurate", "Overestimators"))
  )

h2a_interaction <- lm_robust(
  k_outgroup_ingroup ~ condition * estimation_group,
  se_type = "HC2",
  data = model_data_wide
)

summary(h2a_interaction)
```
```{r}
format_p <- function(p) ifelse(p < .001, "< .001", sprintf("%.3f", p))
add_stars <- function(p) case_when(p < .001 ~ "***", p < .01 ~ "**", p < .05 ~ "*", TRUE ~ "")

h2a_interaction_table <- tidy(h2a_interaction, conf.int = TRUE) |>
  mutate(
    term = term |>
      str_replace("\\(Intercept\\)", "Intercept (Underestimators, Control)") |>
      str_replace("^conditionTreatment$", "Treatment vs. Control") |>
      str_replace("^estimation_groupAccurate$", "Accurate vs. Underestimators") |>
      str_replace("^estimation_groupOverestimators$", "Overestimators vs. Underestimators") |>
      str_replace("conditionTreatment:estimation_groupAccurate", 
                  "Treatment vs. Control × Accurate vs. Underestimators") |>
      str_replace("conditionTreatment:estimation_groupOverestimators", 
                  "Treatment vs. Control × Overestimators vs. Underestimators")
  ) |>
  transmute(
    Predictor = term,
    Estimate = paste0(sprintf("%.3f", estimate), add_stars(p.value)),
    SE = sprintf("%.3f", std.error),
    `95% CI Lower` = sprintf("%.3f", conf.low),
    `95% CI Higher` = sprintf("%.3f", conf.high),
    p = format_p(p.value)
  )

h2a_interaction_table |>
  tt() |>
  save_tt(here(table_dir, "treatment_effect_interaction.docx"), overwrite = TRUE)

h2a_interaction_table |> tt()
```


```{r}
# Predicted means at each cell (condition × estimation_group)
h2a_means <- avg_predictions(
  h2a_interaction,
  by = c("estimation_group", "condition")
) |>
  as_tibble()

h2a_means |> tt()

# Treatment effect within each estimation group
h2a_simple_effects <- avg_comparisons(
  h2a_interaction,
  variables = "condition",
  by = "estimation_group"
) |>
  as_tibble()

h2a_simple_effects |> tt()
```

Check for significant backfire effect:
```{r}
h2a_nonunder <- lm_robust(
  k_outgroup_ingroup ~ condition,
  se_type = "HC2",
  data = model_data_wide |> filter(deviation >= 0)
)

summary(h2a_nonunder)
```

BayesFactors:
```{r}
# JZS priors at three standard widths
bf_medium    <- ttestBF(formula = k_outgroup_ingroup ~ condition, 
                         data = model_data_wide |> filter(deviation >= 0), rscale = "medium")     # r = sqrt(2)/2 ≈ 0.707
bf_wide      <- ttestBF(formula = k_outgroup_ingroup ~ condition, 
                         data = model_data_wide |> filter(deviation >= 0), rscale = "wide")       # r = 1
bf_ultrawide <- ttestBF(formula = k_outgroup_ingroup ~ condition, 
                         data = model_data_wide |> filter(deviation >= 0), rscale = "ultrawide")  # r = sqrt(2)

bf_medium
bf_wide
bf_ultrawide
```
```{r}
bf_extract <- function(bf_obj, prior_label, target = "interaction") {
  raw <- extractBF(bf_obj)
  tibble(
    prior    = prior_label,
    BF10     = raw$bf,
    BF01     = 1 / BF10,
    log_BF10 = log(BF10),
    mc_error = raw$error,
    evidence = case_when(
      BF01 > 10 ~ paste("Strong evidence for null",   target),
      BF01 > 3  ~ paste("Moderate evidence for null", target),
      BF01 > 1  ~ paste("Weak evidence for null",     target),
      BF10 > 10 ~ paste("Strong evidence for",        target),
      BF10 > 3  ~ paste("Moderate evidence for",      target),
      TRUE      ~ paste("Weak evidence for",          target)
    )
  )
}

bf_subgroup_summary <- bind_rows(
  bf_extract(bf_medium,    "medium",    target = "effect"),
  bf_extract(bf_wide,      "wide",      target = "effect"),
  bf_extract(bf_ultrawide, "ultrawide", target = "effect")
)
bf_subgroup_summary |> tt()
```

## Exploratory: Non-Linear Interaction

```{r}
model_data_wide_dev <- model_data_wide_dev |>
  mutate(condition_ord = ordered(condition, levels = c("Control", "Treatment")))

f4_gam <- gam(
  k_outgroup_ingroup ~ condition +
    s(affpol_weighted, bs = "tp") +
    s(affpol_weighted, by = condition_ord, bs = "tp"),
  data   = model_data_wide_dev,
  method = "REML"
)
summary(f4_gam)
```
```{r}
#| fig-width: 4
#| fig-height: 3
#| fig-dpi: 600
pred_grid <- predictions(
  f4_gam,
  newdata = datagrid(
    affpol_weighted = seq(min(model_data_wide_dev$affpol_weighted),
                          max(model_data_wide_dev$affpol_weighted),
                          length.out = 200),
    condition = c("Control", "Treatment")
  )
  ) |> 
  as_tibble()

f4_gam_plot <- ggplot(pred_grid,
                     aes(x = affpol_weighted, y = estimate,
                         color = condition, fill = condition)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.2) +
  geom_rug(data = model_data_wide_dev,
           aes(x = affpol_weighted, color = condition),
           inherit.aes = FALSE, alpha = 0.3, sides = "b") +
  scale_fill_manual(values = c("#fdae61", "#abd9e9")) +
  scale_color_manual(values = c("#f46d43", "#74add1")) +
  labs(
    x = "Affective Polarization (z-scored)",
    y = expression(italic(K)[outgroup] - italic(K)[ingroup]),
    color = NULL, fill = NULL
  ) +
  theme(legend.position = "top")

f4_gam_plot

ggsave(
  here(fig_dir, "GAM_smooths.png"),
  height = 3,
  width  = 4,
  dpi    = 600
)
```

## Exploratory: Bayesian Model Comparison

```{r}
f4_null_med <- lmBF(
  k_outgroup_ingroup ~ condition + affpol_weighted,
  data = model_data_wide_dev, rscaleFixed = "medium", rscaleCont = "medium"
)
f4_inter_med <- lmBF(
  k_outgroup_ingroup ~ condition + affpol_weighted + condition:affpol_weighted,
  data = model_data_wide_dev, rscaleFixed = "medium", rscaleCont = "medium"
)

f4_null_wid <- lmBF(
  k_outgroup_ingroup ~ condition + affpol_weighted,
  data = model_data_wide_dev, rscaleFixed = "wide", rscaleCont = "wide"
)
f4_inter_wid <- lmBF(
  k_outgroup_ingroup ~ condition + affpol_weighted + condition:affpol_weighted,
  data = model_data_wide_dev, rscaleFixed = "wide", rscaleCont = "wide" 
)

f4_null_ult <- lmBF(
  k_outgroup_ingroup ~ condition + affpol_weighted,
  data = model_data_wide_dev, rscaleFixed = "ultrawide", rscaleCont = "ultrawide"
)
f4_inter_ult <- lmBF(
  k_outgroup_ingroup ~ condition + affpol_weighted + condition:affpol_weighted,
  data = model_data_wide_dev, rscaleFixed = "ultrawide", rscaleCont = "ultrawide"
)

# Compute the BFs
f4_bf_med <- f4_inter_med / f4_null_med
f4_bf_wid <- f4_inter_wid / f4_null_wid
f4_bf_ult <- f4_inter_ult / f4_null_ult

f4_bf_med <- recompute(f4_bf_med, iterations = 100000)
f4_bf_wid <- recompute(f4_bf_wid, iterations = 100000)
f4_bf_ult <- recompute(f4_bf_ult, iterations = 100000)
```
```{r}
extract_rscale <- function(bf_obj) {
  prior <- bf_obj@numerator[[1]]@prior$rscale
  tibble(
    r_fixed = prior$fixed,
    r_cont  = prior$continuous
  )
}

f4_table <- bind_rows(
  bind_cols(extract_rscale(f4_inter_med), bf_extract(f4_bf_med, "medium")),
  bind_cols(extract_rscale(f4_inter_wid), bf_extract(f4_bf_wid, "wide")),
  bind_cols(extract_rscale(f4_inter_ult), bf_extract(f4_bf_ult, "ultrawide"))
) |>
  transmute(
    Prior             = str_to_title(prior),
    `r (Categorical)` = sprintf("%.3f", r_fixed),
    `r (Continuous)`  = sprintf("%.3f", r_cont),
    `BF₁₀`            = sprintf("%.3f", BF10),
    `BF₀₁`            = sprintf("%.2f", BF01),
    `MC Error`        = sprintf("%.2f", mc_error * 100)
  )

f4_table |> tt()

f4_table |> tt() |> save_tt(here(table_dir, "Bayesian_model_comp.docx"), overwrite = TRUE)
```

```{r}
# Sample from the posterior of the medium interaction model
f4_posterior <- posterior(f4_inter_med, iterations = 10000)

summary(f4_posterior)
```

### Non-Registered Replications: Partisan Differences

```{r}
party1 <- lm_robust(k_outgroup_ingroup ~ party + rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data_wide)

summary(party1)
```

```{r}
party1.table <- modelsummary(party1, 
                             shape = term ~ model + statistic,
                             title = 'In-Group Favouritism ~ Party',
                             coef_rename = c("partyRepublican" = "Republican - Democrat"),
                             coef_omit = c(1, 3:14),
                             estimate  = "{estimate}{stars}",
                             statistic = c("{std.error}",
                                           "conf.int",
                                           "p.value"),
                             conf_level = .95,
                             stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                             fmt = 3,
                             notes = list('p-values: * < .05, ** < .01, *** < .001'),
                             gof_omit = ".*")

party1.table

party1.table |> save_tt(here(table_dir, "party1.table.docx"), overwrite = T)
```












