---
title: "Democratic Effort Power Analyses"
format:
  html:
    theme: paper
    code-tools: true
    code-fold: true
    toc: true
    toc-fold: false
    toc-location: right
    toc-depth: 3
    toc-title: "Contents"
embed-resources: true
warning: false
---

```{r}
#| include: false
# set error messages to English
Sys.setenv(LANG = "en")

knitr::opts_chunk$set(
  fig.align = "center", 
  fig.retina = 3,
  fig.width = 6, 
  fig.height = (6 * 0.618),
  out.width = "90%", 
  collapse = FALSE
)

options(
  digits = 4,
  width = 300,
  dplyr.summarise.inform = FALSE)
```

```{r}
#| include: false
#| warning: true
#| message: true
if (!require("groundhog", quietly = TRUE)) {
  install.packages("groundhog")
}
library("groundhog")

# use groundhog to install and load packages
pkgs <- c("here",         # System path management
          "tidyverse",    # ggplot, dplyr, %>%, and friends
          "DeclareDesign",    # Tools to plan, implement, analyze empirical research
          "modelsummary", # Data and model summaries with tables and plots
          "tinytable",    # Lightweight package to create tables
          "afex",         # Interface for (general) linear mixed models
          "marginaleffects",      # Compute marginal means, slopes, predictions, etc.
          "ggdist",       # Additional plot elements
          "patchwork",    # Combine multiple ggplots
          "showtext",     # Additional fonts for plots etc
          "tictoc"        # Timing functions
          )

groundhog.library(pkgs, "2026-01-01")
```

```{r}
#| include: false

showtext_auto()
showtext_opts(dpi = 600) # change to 96 on Windows

font_add_google("Public Sans", "Public Sans")
font_add_google("Public Sans", "Public Sans Light", regular.wt = 300)
font_add_google("Public Sans", "Public Sans Semibold", regular.wt = 600)

# Custom ggplot theme
theme_nice <- function(base_size = 14) {
  # General text family and base size
  theme_minimal(base_family = "Public Sans", base_size = base_size) +
    theme(
      # Remove minor grid lines for cleaner look
      panel.grid.minor = element_blank(),
      # Customize major grid lines color
      panel.grid.major = element_blank(),
      # Plot title: bold and custom relative size
      plot.title = element_text(family = "Public Sans", face = "bold", size = rel(1.1), margin = margin(b = 10)),
      # Subtitle and caption: lighter weight 
      plot.subtitle = element_text(family = "Public Sans Light", face = "plain"),
      plot.caption = element_text(family = "Public Sans Light", face = "plain"),
      # Axis text: Relative size
      axis.text = element_text(size = rel(0.9)),
      # Axis titles: Semibold and custom relative size
      axis.title = element_text(family = "Public Sans Semibold", size = rel(0.9)),
      axis.title.x = element_text(hjust = 0.5),
      axis.title.y = element_text(hjust = 0.5),
      # Axis lines
      axis.line = element_line(color = "grey25", linewidth = 0.8),
      # Facet labels: bold text on grey background
      strip.text = element_text(
        family = "Public Sans Semibold", face = "plain",
        size = rel(1),
        hjust = 0.5
      ),
      strip.background = element_rect(fill = "grey95", color = "grey25", linewidth = 0.8), 
      # Legend styling: compact and top-positioned
      legend.title = element_text(family = "Public Sans Semibold", size = rel(0.8)),
      legend.text = element_text(size = rel(0.8)),
      legend.position = "top",
      legend.justification = "center",
      legend.title.position = "left",
      legend.margin = margin(l = 0, t = 0)
    )
}

theme_set(theme_nice())
update_geom_defaults("text", list(family = "Public Sans"))
update_geom_defaults("label", list(family = "Public Sans"))
```

# Democratic Effort Task Power Analyses 

We want to simulate data to calculate the power to reliably detect within-subject differences of Cohens D = 0.20 in k between the in-group and the out-group conditions. 

## 1. Power to detect k_ingroup - k_outgroup given the smallest effect size of interest of Cohens D = 0.20

### Setup 

Declare the model
```{r}
# Parameters
N <- 100  # Sample size
cohens_d <- 0.20  # Smallest effect size of interest
sd_diff <- 1  # SD of difference scores (standardized)
mean_diff <- cohens_d * sd_diff  # Mean difference = 0.20

# Declare model: within-subject design
model <- declare_model(
  N = N,
  # Generate individual difference scores (k_ingroup - k_outgroup)
  # Mean = 0.20, SD = 1 gives Cohen's d = 0.20
  diff = rnorm(N, mean = mean_diff, sd = sd_diff),
  
  # Generate k_ingroup (arbitrary mean/SD)
  k_ingroup = rnorm(N, mean = 0, sd = 1),
  
  # Calculate k_outgroup so difference is as specified
  k_outgroup = k_ingroup - diff
)

```

Declare the inquiry (estimate)
```{r}
inquiry <- declare_inquiry(
  ATE = mean(k_ingroup - k_outgroup)
)
```

Declare the estimator (paired t-test)
```{r}
estimator <- declare_estimator(
  handler = label_estimator(function(data) {
    test <- t.test(data$k_ingroup, data$k_outgroup, paired = TRUE)
    
    tibble(
      estimate = test$estimate,
      statistic = test$statistic,
      p.value = test$p.value,
      conf.low = test$conf.int[1],
      conf.high = test$conf.int[2]
    )
  }),
  inquiry = "ATE",
  label = "paired_t_test"
)
```

Declare the diagnosands
```{r}
diagnosands <- declare_diagnosands(
  mean_estimate = mean(estimate),
  power = mean(p.value < 0.05),
  ci_lower = mean(conf.low, na.rm = TRUE),
  ci_upper = mean(conf.high, na.rm = TRUE)
)
```

Put the design together
```{r}
design_diff = model + inquiry + estimator

# Test simulation
diagnose_design(design_diff, 
                sims = 1000,
                diagnosands = diagnosands)
```

### Simulation

```{r}
#| warning: false
#| message: false
designs_list_diff <- redesign(design_diff,
                         N = seq(50, 300, by = 50)
   )

n_sim = 5000

cat("Running", length(designs_list_diff), "designs with", n_sim, "simulations each...\n")
cat("Total simulations:", length(designs_list_diff) * n_sim, "\n")

tic("Sensitivity analysis")
diagnosis_diff <- diagnose_design(designs_list_diff, 
                                  sims = n_sim,
                                  diagnosands = diagnosands)
toc()
```

### Visualisation

```{r}
#| fig-height: 4 
#| fig-width: 5.5 
#| fig-dpi: 600
fig_diff_data <- diagnosis_diff$diagnosands_df |> 
  rename("Cohen's d" = mean_estimate,
         "SE(Cohen's d)" = `se(mean_estimate)`)

fig_diff <- fig_diff_data |>
  ggplot(aes(x = N, y = power)) +
  geom_line(color = "firebrick3", 
            linewidth = 1) +
  geom_point(color = "firebrick3",
             size = 2) +
  geom_hline(yintercept = 0.90,
             linetype = "dashed",
             color = "black",
             linewidth = 0.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0.2, 1),
                     breaks = seq(0, 1, by = 0.1)) +
  scale_x_continuous(breaks = seq(50, 300, by = 50)) +
  labs(title = "Power to detect Cohen's d = 0.20",
       x = "Sample size (N)",
       y = "Power") +
  theme_nice() +
  theme(legend.position = "none",
      # panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.line = element_line(color = "grey25", linewidth = 0.6),
      axis.ticks = element_line(color = "grey25", linewidth = 0.6),
      axis.ticks.length = unit(0.2, "cm"))

fig_diff
```

```{r}
diff_sim_table <- fig_diff_data |> 
  transmute("Sample size (N)" = N,
            "Cohen's d" = round(`Cohen's d`, 2),
            "Power" = power,
            "SE" = `se(power)`) |> 
  tt(theme = "striped",
     notes = c("Power diagnosis based on 5000 simulations. Power estimates with bootstrapped standard errors."))

diff_sim_table
```

## 2. Power to detect correlation between k_all and between-subject variables

### Setup

Declare the model
```{r}
# Parameters
N <- 100  # Sample size
r <- 0.20  # Correlation coefficient

model_corr <- declare_model(
  N = N,
  X = rnorm(N, mean = 0, sd = 1),  # Between-subject predictor
  k_all = correlate(given = X, rho = r, rnorm, mean = 0, sd = 1)  # Correlated k_all
)

# Declare the inquiry (true correlation)
inquiry_corr <- declare_inquiry(
  correlation = r
)
```

Declare the estimator
```{r}
# Declare the estimator (linear regression)
estimator_corr <- declare_estimator(
  handler = label_estimator(function(data) {
    model <- lm(k_all ~ X, data = data)
    coef_summary <- summary(model)$coefficients
    
    tibble(
      estimate = coef_summary["X", "Estimate"],
      statistic = coef_summary["X", "t value"],
      p.value = coef_summary["X", "Pr(>|t|)"],
      se = coef_summary["X", "Std. Error"],
      conf.low = estimate - 1.96 * se,
      conf.high = estimate + 1.96 * se
    )
  }),
  inquiry = "correlation",
  label = "lm"
)
```

Declare the diagnosands
```{r}
diagnosands_corr <- declare_diagnosands(
  mean_estimate = mean(estimate),
  power = mean(p.value < 0.05),
  ci_lower = mean(conf.low, na.rm = TRUE),
  ci_upper = mean(conf.high, na.rm = TRUE)
)
```

Put the design together
```{r}
design_corr <- model_corr + inquiry_corr + estimator_corr

# Test simulation
diagnose_design(design_corr, 
                sims = 1000,
                diagnosands = diagnosands_corr)
```

### Simulation

```{r}
#| warning: false
#| message: false

designs_list_corr <- redesign(design_corr,
                              N = seq(50, 300, by = 50),
                              r = seq(0.10, 0.2, by = 0.1))

n_sim <- 5000
cat("Running", length(designs_list_corr), "designs with", n_sim, "simulations each...\n")
cat("Total simulations:", length(designs_list_corr) * n_sim, "\n")

tic("Correlation power analysis")
diagnosis_corr <- diagnose_design(designs_list_corr, 
                                  sims = n_sim,
                                  diagnosands = diagnosands_corr)
toc()
```

```{r}
#| fig-height: 4 
#| fig-width: 5.5 
#| fig-dpi: 600
fig_corr_data <- diagnosis_corr$diagnosands_df |> 
  rename("Correlation (r)" = mean_estimate,
         "SE(r)" = `se(mean_estimate)`)

fig_corr <- fig_corr_data |>
  ggplot(aes(x = N, y = power)) +
  geom_line(color = "firebrick3", 
            linewidth = 1) +
  geom_point(color = "firebrick3",
             size = 2) +
  geom_hline(yintercept = 0.90,
             linetype = "dashed",
             color = "black",
             linewidth = 0.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1),
                     breaks = seq(0, 1, by = 0.2)) +
  scale_x_continuous(breaks = seq(50, 300, by = 50)) +
  labs(title = "Power to detect correlations with k_all",
       x = "Sample size (N)",
       y = "Power") +
  facet_wrap(~r, labeller = labeller(r = function(x) paste0("r = ", x))) +
  theme_nice() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        axis.line = element_line(color = "grey25", linewidth = 0.6),
        axis.ticks = element_line(color = "grey25", linewidth = 0.6),
        axis.ticks.length = unit(0.2, "cm"))

fig_corr
```

```{r}
corr_sim_table <- fig_corr_data |> 
  transmute("Sample size (N)" = N,
            "Correlation (r)" = r,
            "Estimated r" = round(`Correlation (r)`, 2),
            "Power" = power,
            "SE" = `se(power)`) |> 
  tt(theme = "striped",
     notes = c("Power diagnosis based on 5000 simulations. Power estimates with bootstrapped standard errors."))

corr_sim_table
```

```{r}
#| fig-height: 8 
#| fig-width: 6 
#| fig-dpi: 600
final_fig <- fig_diff / fig_corr + 
  plot_layout(heights = c(1, 1.2)) +
  plot_annotation(tag_levels = 'A')

# ggsave(here(sim_dir, "simulation_results.png"), 
#        plot = final_fig,
#        height = 10, 
#        width = 9, 
#        dpi = 600)

final_fig
```







