---
title: "Democratic Effort Study 1: Logistic Regressions"
format:
  html:
    theme: paper
    code-tools: true
    code-fold: true
    toc: true
    toc-fold: false
    toc-location: right
    toc-depth: 3
    toc-title: "Contents"
embed-resources: true
warning: false
---

```{r}
#| include: false
# set error messages to English
Sys.setenv(LANG = "en")

knitr::opts_chunk$set(
  fig.align = "center", 
  fig.retina = 3,
  fig.width = 6, 
  fig.height = (6 * 0.618),
  out.width = "90%", 
  collapse = FALSE
)

options(
  digits = 4,
  width = 300,
  dplyr.summarise.inform = FALSE)
```

```{r}
#| include: false
#| warning: true
#| message: true
if (!require("groundhog", quietly = TRUE)) {
  install.packages("groundhog")
}
library("groundhog")

# use groundhog to install and load packages
pkgs <- c("here",         # System path management
          "tidyverse",    # ggplot, dplyr, %>%, and friends
          "modelsummary", # Data and model summaries with tables and plots
          "tinytable",    # Lightweight package to create tables
          "afex",         # Interface for (general) linear mixed models
          "broom.mixed",  # Get tidy model objects
          "marginaleffects"  # Compute marginal means, slopes, predictions, etc.
          )

groundhog.library(pkgs, "2026-01-01")
```

```{r}
#| include: false
data_dir <- here("01_data", "final") 
table_dir <- here("02_output", "tables", "log_regressions") 

dir.create(table_dir, recursive = TRUE, showWarnings = TRUE)
```

# Democratic Effort Task Regressions

## Load and prepare data

```{r}
load(here(data_dir, "demeffort_survey_data.RData")) 
load(here(data_dir, "demeffort_combined_data.RData"))
```

## Generalised linear mixed models

Modelling strategy:

1.  Fix maximum models (Barr et al., 2013)

2.  If models don't converge:

    1.  Remove the correlations between random slope and intercept

    2.  Remove interactions between random slope effects

    3.  Remove random slope effects with a variance close to 0

    4.  Run models only including a random intercept

```{r}
model_data <- combined_data |> 
  mutate(rew = factor(rew, levels = c(4, 12, 20))) |> 
  filter(!is.na(decision))
```

### Model 1: Full in-group / out-group / all model

```{r}
# Error: number of observations (=7161) < number of random effects (=7668)
# m1.full <- mixed(decision ~ benefit * eff * rew + (1 + benefit * eff * rew | PID), 
#             data = model_data, 
#             family = binomial, 
#             method = "LRT",
#             control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e5)), 
#             return = "merMod", 
#             expand_re = T) 
# 
# summary(m1.full)
```

```{r}
# Model failed to converge: degenerate  Hessian with 2 negative eigenvalues
# m1.nocorr <- mixed(decision ~ benefit * eff * rew + (1 + benefit * eff * rew || PID),
#             data = model_data,
#             family = binomial,
#             method = "LRT",
#             control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e5)),
#             return = "merMod",
#             expand_re = T)
# 
# summary(m1.nocorr)
```

```{r}
# Model failed to converge with max|grad| = 0.347206 (tol = 0.002, component 1)
# m1.noint <- mixed(decision ~ benefit * eff * rew + (1 + benefit + eff + rew || PID),
#             data = model_data,
#             family = binomial,
#             method = "LRT",
#             control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e5)),
#             return = "merMod",
#             expand_re = T)
# 
# summary(m1.noint)
```

```{r}
m1 <- mixed(decision ~ benefit * eff * rew + (1 + benefit | PID), 
            data = model_data, 
            family = binomial, 
            method = "LRT",
            control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e5)), 
            return = "mixed", 
            expand_re = T) 
```

```{r}
anova_table <- m1$anova_table |>
  as_tibble(rownames = "Effect") |>
  select(
    Effect,
    df   = `Chi Df`,
    Chi2 = Chisq,
    p    = `Pr(>Chisq)`
  ) |>
  mutate(
    Effect = Effect |>
      str_replace_all("benefit", "Beneficiary") |>
      str_replace_all("rew",     "Reward") |>
      str_replace_all("eff",     "Effort"),
    Chi2 = sprintf("%.2f", Chi2),
    p    = ifelse(p < .001, "< .001", sprintf("%.3f", p))
  )

anova_table |>
  tt()

anova_table |>
  tt() |>
  save_tt(here(table_dir, "anova_table.docx"), overwrite = TRUE)
```

```{r}
contrasts(m1$data$benefit)
contrasts(m1$data$eff)
contrasts(m1$data$rew)
```


```{r}
m1.mer <- m1$full_model

relabel_terms <- function(x) {
  x |>
    str_replace_all("benefit1", "Everyone") |>
    str_replace_all("benefit2", "Ingroup") |>
    str_replace_all("eff1",     "Effort Level 1") |>
    str_replace_all("eff2",     "Effort Level 2") |>
    str_replace_all("rew1",     "4 Credits") |>
    str_replace_all("rew2",     "12 Credits") |>
    str_replace_all(":",        " \u00d7 ")
}

fe <- tidy(m1.mer, effects = "fixed", conf.int = TRUE) |>
  mutate(
    term = relabel_terms(term),
    term = case_when(
      term == "(Intercept)"                              ~ term,
      str_detect(term, "\u00d7")                         ~ term,
      term %in% c("Everyone", "Ingroup")                 ~ paste0("Beneficiary: ", term),
      term %in% c("Effort Level 1", "Effort Level 2")    ~ paste0("Effort: ", term),
      term %in% c("4 Credits", "12 Credits")             ~ paste0("Reward: ", term),
      TRUE                                               ~ term
    )
  ) |>
  transmute(
    Term     = term,
    Estimate = sprintf("%.3f", estimate),
    SE       = sprintf("%.3f", std.error),
    z        = sprintf("%.2f", statistic),
    `95% CI` = paste0("[", sprintf("%.3f", conf.low), ", ", sprintf("%.3f", conf.high), "]"),
    p        = ifelse(p.value < .001, "< .001", sprintf("%.3f", p.value))
  )

re <- tidy(m1.mer, effects = "ran_pars") |>
  mutate(
    term = term |>
      str_replace_all("re1\\.benefit1", "Everyone") |>
      str_replace_all("re1\\.benefit2", "Ingroup") |>
      str_replace_all("sd__\\(Intercept\\)",                     "SD (Intercept)") |>
      str_replace_all("sd__Everyone",                            "SD (Everyone)") |>
      str_replace_all("sd__Ingroup",                             "SD (Ingroup)") |>
      str_replace_all("cor__\\(Intercept\\)\\.Everyone",         "Cor (Intercept, Everyone)") |>
      str_replace_all("cor__\\(Intercept\\)\\.Ingroup",          "Cor (Intercept, Ingroup)") |>
      str_replace_all("cor__Everyone\\.Ingroup",                 "Cor (Everyone, Ingroup)")
  ) |>
  transmute(
    Term     = term,
    Estimate = sprintf("%.3f", estimate),
    SE       = "",
    z        = "",
    `95% CI` = "",
    p        = ""
  )

n_fe <- nrow(fe)
lmm_table <- bind_rows(fe, re) |>
  tt(notes = "Sum-to-zero (deviation) contrast coding applied to all factors; coefficients represent deviation from the grand mean. Implicit reference levels (coded as the negative sum of displayed contrasts): Beneficiary = Out-group; Effort = Level 3; Reward = 20 Credits.") |>
  group_tt(
    i = list(
      "Fixed Effects"                = 1,
      "Random Effects (Participant)" = n_fe + 1
    )
  ) |>
  style_tt(i = c(1, n_fe + 2), bold = TRUE)

lmm_table

lmm_table |>
  save_tt(here(table_dir, "lmm_table.docx"), overwrite = TRUE)
```

### Model-based predictions

```{r}
m1.ben.means <- avg_predictions(m1.mer, 
                                variables = "benefit",
                                re.form=NA) |> 
  as_tibble() |> 
  select(Condition = benefit, 
         Estimate = estimate, 
         SE = std.error,
         "t" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high)

m1.eff.means <- avg_predictions(m1.mer, 
                                variables = "eff",
                                re.form=NA) |> 
  as_tibble() |> 
  select(Condition = eff, 
         Estimate = estimate, 
         SE = std.error,
         "t" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high) 

m1.rew.means <- avg_predictions(m1.mer, 
                                variables = "rew",
                                re.form=NA) |> 
  as_tibble() |> 
  select(Condition = rew, 
         Estimate = estimate, 
         SE = std.error,
         "t" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high)


m1.means <- bind_rows(m1.ben.means, m1.eff.means, m1.rew.means) |> 
  mutate(
    across(where(is.numeric) & !matches("p"), ~ round(.x, 3)),  # Round all numeric except p
    p = ifelse(round(p, 3) == 0, "< .001", as.character(round(p, 3)))  # Special handling for p
  )

m1.means |> 
  tt() 

m1.means |> 
  tt() |> 
  save_tt(here(table_dir, "model_means.docx"), overwrite = TRUE)
```

### Model-based comparisons

#### Main pre-registered quantities of interest

```{r}
m1.comparisons <- avg_comparisons(m1.mer, 
                                  variables = list(benefit = "pairwise",
                                                   eff = "pairwise",
                                                   rew = "pairwise"),
                                  re.form=NA) |> 
  as_tibble() |> 
  select(Contrast = contrast, 
         Estimate = estimate, 
         SE = std.error,
         "t" = statistic,
         "p" = p.value,
         "95% CI Low" = conf.low,
         "95% CI High" = conf.high) |> 
  mutate(
    across(where(is.numeric) & !matches("p"), ~ round(.x, 3)),  # Round all numeric except p
    p = ifelse(round(p, 3) == 0, "< .001", as.character(round(p, 3)))  # Special handling for p
  )

m1.comparisons |> 
  tt() 

m1.comparisons |> 
  tt() |> 
  save_tt(here(table_dir, "model_comparisons.docx"), overwrite = TRUE)
```


