---
title: "Democratic Effort k-Regressions"
format:
  html:
    theme: paper
    code-tools: true
    code-fold: true
    toc: true
    toc-fold: false
    toc-location: right
    toc-depth: 3
    toc-title: "Contents"
embed-resources: true
warning: false
---

```{r}
#| include: false
# set error messages to English
Sys.setenv(LANG = "en")

knitr::opts_chunk$set(
  fig.align = "center", 
  fig.retina = 3,
  fig.width = 6, 
  fig.height = (6 * 0.618),
  out.width = "90%", 
  collapse = FALSE
)

options(
  digits = 4,
  width = 300,
  dplyr.summarise.inform = FALSE)
```

```{r}
#| include: false
#| warning: true
#| message: true
if (!require("groundhog", quietly = TRUE)) {
  install.packages("groundhog")
}
library("groundhog")

# use groundhog to install and load packages
pkgs <- c("here",         # System path management
          "datawizard",   # Data preparations (e.g., standardise)
          "tidyverse",    # ggplot, dplyr, %>%, and friends
          "estimatr",     # Fast estimators with robust standard errors
          "effectsize",   # Effect sizes: Cohen's D and friends
          "correlation",  # For Bayesian correlations
          "modelsummary", # Data and model summaries with tables and plots
          "tinytable",    # Lightweight package to create tables
          "marginaleffects",      # Compute marginal means, slopes, predictions, etc.
          "broom",        # Tidy up models
          "ggdist",       # Additional plot elements
          "GGally",       # ggpairs and other ggplot2 friends
          "patchwork",    # Combine multiple ggplots
          "showtext"      # Additional fonts for plots etc
          )

groundhog.library(pkgs, "2026-01-01")
```

```{r}
#| include: false
showtext_auto()
showtext_opts(dpi = 600) # change to 96 on Windows

font_add_google("Public Sans", "Public Sans")
font_add_google("Public Sans", "Public Sans Light", regular.wt = 300)
font_add_google("Public Sans", "Public Sans Semibold", regular.wt = 600)

# Custom ggplot theme
theme_nice <- function(base_size = 14) {
  # General text family and base size
  theme_minimal(base_family = "Public Sans", base_size = base_size) +
    theme(
      # Remove minor grid lines for cleaner look
      panel.grid.minor = element_blank(),
      # Customize major grid lines color
      panel.grid.major = element_blank(),
      # Plot title: bold and custom relative size
      plot.title = element_text(family = "Public Sans", face = "bold", size = rel(1.1), margin = margin(b = 10)),
      # Subtitle and caption: lighter weight 
      plot.subtitle = element_text(family = "Public Sans Light", face = "plain"),
      plot.caption = element_text(family = "Public Sans Light", face = "plain"),
      # Axis text: Relative size
      axis.text = element_text(size = rel(0.9)),
      # Axis titles: Semibold and custom relative size
      axis.title = element_text(family = "Public Sans Semibold", size = rel(0.9)),
      axis.title.x = element_text(hjust = 0.5),
      axis.title.y = element_text(hjust = 0.5),
      # Axis lines
      axis.line = element_line(color = "grey25", linewidth = 0.8),
      # Facet labels: bold text on grey background
      strip.text = element_text(
        family = "Public Sans Semibold", face = "plain",
        size = rel(1),
        hjust = 0.5
      ),
      strip.background = element_rect(fill = "grey95", color = "grey25", linewidth = 0.8), 
      # Legend styling: compact and top-positioned
      legend.title = element_text(family = "Public Sans Semibold", size = rel(0.8)),
      legend.text = element_text(size = rel(0.8)),
      legend.position = "top",
      legend.justification = "center",
      legend.title.position = "left",
      legend.margin = margin(l = 0, t = 0)
    )
}

theme_set(theme_nice())
update_geom_defaults("text", list(family = "Public Sans"))
update_geom_defaults("label", list(family = "Public Sans"))

agent_limits <- c("VOTEALL", "VOTERURAL", "VOTEURBAN")
agent_colors <- c("VOTEALL" = "grey10", "VOTERURAL" = "firebrick3", "VOTEURBAN" = "slateblue4")  
agent_fills <- c("VOTEALL" = "grey10", "VOTERURAL" = "firebrick3", "VOTEURBAN" = "slateblue4")   
agent_lines <- c("VOTEALL" = "solid", "VOTERURAL" = "dashed", "VOTEURBAN" = "dashed")

benefit_limits <- c("In-Group", "Everyone", "Out-Group")
benefit_colors <- c("Everyone" = "grey20", "In-Group" = "#009988", "Out-Group" = "#EE7733") 
benefit_fills <- c("Everyone" = "grey20", "In-Group" = "#009988", "Out-Group" = "#EE7733")
benefit_lines <- c("Everyone" = "solid", "In-Group" = "dashed", "Out-Group" = "dashed")
```

```{r}
#| include: false
data_dir <- here("01_data", "final") 
table_dir <- here("02_output", "tables", "k_regressions") 
fig_dir <- here("02_output", "figures", "k_regressions") 

dir.create(table_dir, recursive = TRUE, showWarnings = TRUE)
dir.create(fig_dir, recursive = TRUE, showWarnings = TRUE)
```

# Democratic Effort Task Regressions 

## Load and prepare data

```{r}
load(here(data_dir, "demeffort_survey_data.RData")) 
load(here(data_dir, "demeffort_combined_data.RData"))
```

```{r}
model_data <- survey_data |>
  standardize(
    select = c(k_all, k_rural, k_urban, k_ingroup, k_outgroup, 
               k_mean, k_outgroup_ingroup, k_all_ingroup, k_outgroup_all,
               beta, max_boxed)
  ) |> 
  mutate(gender = fct_collapse(gender,
    "Woman" = c("Woman", "Other") # only one other case, otherwise models don't converge
  )) |> 
  droplevels() |> 
  mutate(donation_decision = factor(donation_decision, levels = c("Keep", "Donate")))
```

## Task Validation

### Preliminary

```{r}
t.test(survey_data$k_outgroup, survey_data$k_ingroup, paired = TRUE) 
cohens_d(survey_data$k_outgroup, survey_data$k_ingroup, paired = TRUE)
```

### H1d

H1d. Participants who show greater motivation / less effort discounting for non-partisan democratic engagement will also be more likely to donate the money to a non-partisan cause after the task rather than accepting it as bonus payment for themselves. 
To investigate H1d, we will run a linear regression model using the motivation / effort discounting parameter for all areas (k_all) as the outcome, and the decision after the task to donate the money (1) or to keep the money (0) as predictor variable: k_all ~ Donation 

```{r}
h1d <- lm_robust(k_all ~ donation_decision, se_type = "HC2", data = model_data)
summary(h1d)
```

```{r}
h1d.table <- modelsummary(h1d, 
                          shape = term ~ model + statistic,
                          title = 'Table H1d: k_all ~ Donation Decision',
                          coef_rename = c("donation_decisionDonate" = "Donate - Keep"),
                          coef_omit = 1,
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h1d.table

h1d.table |> save_tt(here(table_dir, "h1d.table.docx"), overwrite = T)
```

## Partisan Differences

### H2a

H2a. Republicans will show greater motivation / less effort discounting for rural areas than Democrats. 
To investigate H2a, we will run a linear regression model using  the motivation / effort discounting parameter for rural areas (k_rural) as outcome variable and partisanship (Democrat / Republican) as predictor variable: k_rural ~ Partisanship + Controls

```{r}
h2a <- lm_robust(k_rural ~ party + rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data)
h2a_nocon <- lm_robust(k_rural ~ party, se_type = "HC2", data = model_data)

summary(h2a)
```

```{r}
h2a.models <- list(
  "Pre-Registered" = h2a, 
  "Without Controls" = h2a_nocon)

h2a.table <- modelsummary(h2a.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H2a: k_rural ~ Party',
                          coef_rename = c("partyRepublican" = "Republican - Democrat"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h2a.table

h2a.table |> save_tt(here(table_dir, "h2a.table.docx"), overwrite = T)
```

### H2b

H2b. Democrats will show greater motivation / less effort discounting for urban areas than Republicans. 
To investigate H2b, we will run a linear regression model using  the motivation / effort discounting parameter for urban areas (k_urban) as outcome variable and partisanship (Democrat / Republican) as predictor variable: k_urban ~ Partisanship + Controls

```{r}
h2b <- lm_robust(k_urban ~ party + rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data)
h2b_nocon <- lm_robust(k_urban ~ party, se_type = "HC2", data = model_data)

summary(h2b)
```

```{r}
h2b.models <- list(
  "Pre-Registered" = h2b, 
  "Without Controls" = h2b_nocon)

h2b.table <- modelsummary(h2b.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H2b: k_urban ~ Party',
                          coef_rename = c("partyRepublican" = "Republican - Democrat"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h2b.table

h2b.table |> save_tt(here(table_dir, "h2b.table.docx"), overwrite = T)
```

### H2c

H2c: Democrats and Republicans differ in their motivation / effort discounting for non-partisan democratic engagement.
To investigate H2c, we will run a linear regression model using the motivation / effort discounting parameter for all areas (k_all) as outcome variable and partisanship (Democrat / Republican) as predictor variable: k_all ~ Partisanship + Controls

```{r}
h2c <- lm_robust(k_all ~ party + rururb01 + age + gender + education + max_boxes, data = model_data)
h2c_nocon <- lm_robust(k_all ~ party, data = model_data)

summary(h2c)
```

```{r}
h2c.models <- list(
  "Pre-Registered" = h2c, 
  "Without Controls" = h2c_nocon)

h2c.table <- modelsummary(h2c.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H2c: k_all ~ Party',
                          coef_rename = c("partyRepublican" = "Republican - Democrat"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h2c.table

h2c.table |> save_tt(here(table_dir, "h2c.table.docx"), overwrite = T)
```

### H2d

H2d: Democrats and Republicans differ in their in-group favouritism (i.e., the difference between motivation / effort discounting in the in-group - the out-group condition).
To investigate H2d, we will run a linear regression model using the difference between the motivation / effort discounting parameters for the in-group and out-group (k_outgroup - k_ingroup) as the outcome and partisanship (Democrat / Republican) as predictor variable: (k_outgroup - k_ingroup) ~ Partisanship + Controls

```{r}
h2d <- lm_robust(k_outgroup_ingroup ~ party + rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data)
h2d_nocon <- lm_robust(k_outgroup_ingroup ~ party, se_type = "HC2", data = model_data)

summary(h2d)
```

```{r}
h2d.models <- list(
  "Pre-Registered" = h2d, 
  "Without Controls" = h2d_nocon)

h2d.table <- modelsummary(h2d.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H2d: In-Group Favouritism ~ Party',
                          coef_rename = c("partyRepublican" = "Republican - Democrat"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h2d.table

h2d.table |> save_tt(here(table_dir, "h2d.table.docx"), overwrite = T)
```

## Political Correlates

### H3a

H3a. Support for democracy will be positively correlated with greater motivation / less effort discounting for non-partisan democratic engagement.
To investigate H3a, we will run a linear regression model using the motivation / effort discounting parameter for all areas (k_all) as the outcome and support for democracy as predictor variable: k_all ~ Democratic Support + Controls

```{r}
h3a <- lm_robust(k_all ~ democratic_support + rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data)
h3a_nocon <- lm_robust(k_all ~ democratic_support, se_type = "HC2", data = model_data)

summary(h3a)
```

```{r}
h3a.models <- list(
  "Pre-Registered" = h3a, 
  "Without Controls" = h3a_nocon)

h3a.table <- modelsummary(h3a.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H3a: k_all ~ democratic_support',
                          coef_rename = c("democratic_support" = "Support for Democracy"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h3a.table

h3a.table |> save_tt(here(table_dir, "h3a.table.docx"), overwrite = T)
```

#### F3a 

Democratic Support, Item-level

```{r}
f3a <- lm_robust(k_all ~ dem_01_r + dem_02_r + dem_03_r + dem_04 + dem_05_r + dem_06_r + dem_07_r + 
            rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data)
f3a_nocon <- lm_robust(k_all ~ dem_01_r + dem_02_r + dem_03_r + dem_04 + dem_05_r + dem_06_r + dem_07_r, 
                se_type = "HC2", data = model_data)

summary(f3a)
```

```{r}
f3a.models <- list(
  "Pre-Registered" = f3a,
  "Without Controls" = f3a_nocon)

f3a.table <- modelsummary(f3a.models, 
                          shape = term ~ model + statistic,
                          title = 'Table f3a: k_all ~ democratic_support items',
                          coef_rename = c(
                            "dem_01_r" = "Freedom of Expression",
                            "dem_02_r" = "Freedom of Association", 
                            "dem_03_r" = "Universal Suffrage",
                            "dem_04" = "Free and Fair Elections",
                            "dem_05_r" = "Judicial Constraints on Executive",
                            "dem_06_r" = "Legislative Constraints on Executive",
                            "dem_07_r" = "Equality Before the Law"
                          ),
                          coef_omit = c(1, 9:18),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

f3a.table

f3a.table |> save_tt(here(table_dir, "f3a.table.docx"), overwrite = T)
```

### H3b

H3b. Voting importance will be positively correlated with greater motivation / less effort discounting for non-partisan democratic engagement.
To investigate H3b, we will run a linear regression model using the motivation / effort discounting parameter for all areas (k_all) as the outcome and voting importance as predictor variable: k_all ~ Voting Importance + Controls

```{r}
h3b <- lm_robust(k_all ~ vote_importance + rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data)
h3b_nocon <- lm_robust(k_all ~ vote_importance, se_type = "HC2", data = model_data)

summary(h3b)
```

```{r}
h3b.models <- list(
  "Pre-Registered" = h3b, 
  "Without Controls" = h3b_nocon)

h3b.table <- modelsummary(h3b.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H3b: k_all ~ vote_importance',
                          coef_rename = c("vote_importance" = "Voting Importance"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h3b.table

h3b.table |> save_tt(here(table_dir, "h3b.table.docx"), overwrite = T)
```

### H3c

H3c. Partisan intensity will be positively correlated with greater motivation / less effort discounting for the political in-group over the political out-group. 
To investigate H3c, we will run a linear regression model using the difference between the motivation / effort discounting parameters for the in-group and out-group (k_outgroup - k_ingroup) as the outcome and partisan intensity as predictor variable: (k_outgroup - k_ingroup) ~ Partisan Intensity + Controls

```{r}
h3c <- lm_robust(k_outgroup_ingroup ~ party_identity_strength + rururb01 + age + gender + education + max_boxes, 
                 se_type = "HC2", data = model_data)
h3c_nocon <- lm_robust(k_outgroup_ingroup ~ party_identity_strength, se_type = "HC2", data = model_data)

summary(h3c)
```

```{r}
h3c.models <- list(
  "Pre-Registered" = h3c, 
  "Without Controls" = h3c_nocon)

h3c.table <- modelsummary(h3c.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H3c: In-Group Favoritism ~ party_identity_strength',
                          coef_rename = c("party_identity_strength" = "Partisan Identity Strength"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h3c.table

h3c.table |> save_tt(here(table_dir, "h3c.table.docx"), overwrite = T)
```

### H3d

H3d. Affective polarisation will be positively correlated with greater motivation / less effort discounting for the political in-group over the political out-group.
To investigate H3c, we will run a linear regression model using the difference between the motivation / effort discounting parameters for the in-group and out-group (k_outgroup - k_ingroup) as the outcome and affective polarisation as the predictor variable: (k_outgroup - k_ingroup) ~ Affective Polarisation + Controls

```{r}
h3d <- lm_robust(k_outgroup_ingroup ~ affpol_weighted + rururb01 + age + gender + education + max_boxes, 
                 se_type = "HC2", data = model_data)
h3d_nocon <- lm_robust(k_outgroup_ingroup ~ affpol_weighted, 
                       se_type = "HC2", data = model_data)

summary(h3d)
```

```{r}
h3d.models <- list(
  "Pre-Registered" = h3d, 
  "Without Controls" = h3d_nocon)

h3d.table <- modelsummary(h3d.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H3d: In-Group Favoritism ~ affpol_weighted',
                          coef_rename = c("affpol_weighted" = "Affective Polarization"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h3d.table

h3d.table |> save_tt(here(table_dir, "h3d.table.docx"), overwrite = T)
```

### Table

```{r}
format_p <- function(p) ifelse(p < .001, "< .001", sprintf("%.3f", p))
add_stars <- function(p) case_when(p < .001 ~ "***", p < .01 ~ "**", p < .05 ~ "*", TRUE ~ "")

extract_focal <- function(model, focal_term, panel) {
  tidy(model, conf.int = TRUE) |>
    filter(term == focal_term) |>
    transmute(
      panel,
      Predictor = focal_term,
      Estimate  = paste0(sprintf("%.3f", estimate), add_stars(p.value)),
      SE        = sprintf("%.3f", std.error),
      `95% CI Lower`    = sprintf("%.3f", conf.low),
      `95% CI Higher`   = sprintf("%.3f", conf.high),
      p         = format_p(p.value)
    )
}

label_map <- c(
  "democratic_support"      = "Support for Democracy",
  "vote_importance"         = "Voting Importance",
  "party_identity_strength" = "Partisan Identity Strength"
)

s5_rows <- bind_rows(
  extract_focal(h3a,       "democratic_support",      "Pre-Registered"),
  extract_focal(h3a_nocon, "democratic_support",      "Without Controls"),
  extract_focal(h3b,       "vote_importance",         "Pre-Registered"),
  extract_focal(h3b_nocon, "vote_importance",         "Without Controls"),
  extract_focal(h3c,       "party_identity_strength", "Pre-Registered"),
  extract_focal(h3c_nocon, "party_identity_strength", "Without Controls")
) |>
  mutate(
    Predictor = recode(Predictor, !!!label_map),
    Model     = panel
  ) |>
  select(Model, Estimate, SE, `95% CI Lower`, `95% CI Higher`, p)

s5_table <- s5_rows |>
  tt(notes = list(
    "Models control for locality of residence, age, gender, education, and general effort capacity. All continuous variables z-scored.",
    "p-values: * < .05, ** < .01, *** < .001"
  )) |>
  group_tt(
    i = list(
      "Support for Democracy ~ K All"               = 1,
      "Voting Importance ~ K All Areas"                   = 3,
      "Partisan Identity Strength ~ In-Group Favoritism"  = 5
    )
  ) |>
  style_tt(i = c(1, 4, 7), bold = TRUE)

s5_table

s5_table |> save_tt(here(table_dir, "s5_political_correlates.docx"), overwrite = TRUE)
```

```{r}
#| message: false
#| fig-width: 12
#| fig-height: 10
#| fig-dpi: 600
scatter_function <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_point(alpha = 0.2, size = 1, color = "#fd8d3c") +
    geom_smooth(method    = "lm",
                se        = TRUE,
                linewidth = 1.2,
                alpha     = 0.15,
                color     = "black",
                fill      = "black",
                na.rm     = TRUE)
}

political_corr_plot <- survey_data |>
  select(
    "K Out\u2212In" = k_outgroup_ingroup,
    "K All"         = k_all,
    "K In"          = k_ingroup,
    "K Out"         = k_outgroup,
    "DemSupp"       = democratic_support,
    "VoteImp"       = vote_importance,
    "PartStr"       = party_identity_strength
  ) |>
  ggpairs(
    lower    = list(continuous = scatter_function),
    upper    = list(continuous = wrap("cor", stars = TRUE, size = 5)),
    diag     = list(continuous = wrap("densityDiag", alpha = 0.5,
                                      fill = "#fd8d3c", color = "#f16913")),
    progress = FALSE
  ) +
  theme_nice() +
  theme(
    axis.text  = element_text(size = 10),
    strip.text = element_text(size = 11)
  )

political_corr_plot

ggsave(
  here(fig_dir, "political_correlogram.png"),
  height = 10,
  width  = 12,
  dpi    = 600
)
```


## General and Mental-Health

### H4a

H4a. Greater general stress will be negatively correlated with greater motivation / less effort discounting for non-partisan, in-group and out-group democratic engagement. 
To investigate H4a, we will run a linear regression model predicting the average motivation / effort discounting parameter across conditions (k_mean = mean(k_all, k_ingroup, k_outgroup)) as the outcome variable and general stress as the predictor variable: k_mean ~ General Stress + Controls

```{r}
h4a <- lm_robust(k_mean ~ stress_general + rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data)
h4a_nocon <- lm_robust(k_mean ~ stress_general, se_type = "HC2", data = model_data)

summary(h4a)
```

```{r}
h4a.models <- list(
  "Pre-Registered" = h4a, 
  "Without Controls" = h4a_nocon)

h4a.table <- modelsummary(h4a.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H4a: k_mean ~ stress_general',
                          coef_rename = c("stress_general" = "General Stress"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h4a.table

h4a.table |> save_tt(here(table_dir, "h4a.table.docx"), overwrite = T)
```

### H4b

H4b. Greater depression scores will be negatively correlated with greater motivation / less effort discounting for non-partisan, in-group and out-group democratic engagement. 
To investigate H4b, we will run a linear regression model predicting the average motivation / effort discounting parameter across conditions (k_mean = mean(k_all, k_ingroup, k_outgroup)) as the outcome variable and depression score as the predictor variable: k_mean ~ Depression Score + Controls

```{r}
h4b <- lm_robust(k_mean ~ depression + rururb01 + age + gender + education + max_boxes, 
          se_type = "HC2", data = model_data)
h4b_nocon <- lm_robust(k_mean ~ depression, se_type = "HC2", data = model_data)

summary(h4b)
```

```{r}
h4b.models <- list(
  "Pre-Registered" = h4b, 
  "Without Controls" = h4b_nocon)

h4b.table <- modelsummary(h4b.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H4b: k_mean ~ depression',
                          coef_rename = c("depression" = "Depression Score"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h4b.table

h4b.table |> save_tt(here(table_dir, "h4b.table.docx"), overwrite = T)
```

### H4c

H4c. Poor general health will be negatively correlated with greater motivation / less effort discounting for non-partisan, in-group and out-group democratic engagement. 
To investigate H4c, we will run a linear regression model predicting the average motivation / effort discounting parameter across conditions (k_mean = mean(k_all, k_ingroup, k_outgroup)) as the outcome variable and general health as the predictor variable: k_mean ~ General Health + Controls

```{r}
h4c <- lm_robust(k_mean ~ health_general + rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data)
h4c_nocon <- lm_robust(k_mean ~ health_general, se_type = "HC2", data = model_data)

summary(h4c)
```

```{r}
h4c.models <- list(
  "Pre-Registered" = h4c, 
  "Without Controls" = h4c_nocon)

h4c.table <- modelsummary(h4c.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H4c: k_mean ~ health_general',
                          coef_rename = c("health_general" = "General Health"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h4c.table

h4c.table |> save_tt(here(table_dir, "h4c.table.docx"), overwrite = T)
```

### H4d

H4d. There are two competing hypotheses regarding the relationship between costs to health due to politics and motivation / effort discounting for democratic engagement:
- H4d1. Greater costs to health due to politics will be positively associated with motivation / less effort discounting, reflecting engagement-driven effects.
- H4d2. Greater costs to health due to politics will be negatively associated with motivation / less effort discounting, reflecting limitations imposed by poor health.
To investigate H4d, we will run a linear regression model predicting the average motivation / effort discounting parameter across conditions (k_mean = mean(k_all, k_ingroup, k_outgroup)) as the outcome variable and health due to politics as the predictor variable: k_mean ~ Health and politics + Controls

```{r}
h4d <- lm_robust(k_mean ~ health_politics + rururb01 + age + gender + education + max_boxes, se_type = "HC2", data = model_data)
h4d_nocon <- lm_robust(k_mean ~ health_politics, se_type = "HC2", data = model_data)

summary(h4d)
```

```{r}
h4d.models <- list(
  "Pre-Registered" = h4d, 
  "Without Controls" = h4d_nocon)

h4d.table <- modelsummary(h4d.models, 
                          shape = term ~ model + statistic,
                          title = 'Table H4d: k_mean ~ health_politics',
                          coef_rename = c("health_politics" = "Health & Politics"),
                          coef_omit = c(1, 3:12),
                          estimate  = "{estimate}{stars}",
                          statistic = c("{std.error}",
                                        "conf.int",
                                        "p.value"),
                          conf_level = .95,
                          stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
                          fmt = 3,
                          notes = list('p-values: * < .05, ** < .01, *** < .001'),
                          gof_omit = ".*")

h4d.table

h4d.table |> save_tt(here(table_dir, "h4d.table.docx"), overwrite = T)
```

### Table

```{r}
health_label_map <- c(
  "stress_general"  = "General Stress",
  "depression"      = "Depression Score",
  "health_general"  = "General Health",
  "health_politics" = "Health Costs Due to Politics"
)

s6_rows <- bind_rows(
  extract_focal(h4a,       "stress_general",  "Pre-Registered"),
  extract_focal(h4a_nocon, "stress_general",  "Without Controls"),
  extract_focal(h4b,       "depression",      "Pre-Registered"),
  extract_focal(h4b_nocon, "depression",      "Without Controls"),
  extract_focal(h4c,       "health_general",  "Pre-Registered"),
  extract_focal(h4c_nocon, "health_general",  "Without Controls"),
  extract_focal(h4d,       "health_politics", "Pre-Registered"),
  extract_focal(h4d_nocon, "health_politics", "Without Controls")
) |>
  mutate(
    Predictor = recode(Predictor, !!!health_label_map),
    Model     = panel
  ) |>
  select(Model, Estimate, SE, `95% CI Lower`, `95% CI Higher`, p)

s6_table <- s6_rows |>
  tt(notes = list(
    "Models control for locality of residence, age, gender, education, and general effort capacity. All continuous variables z-scored.",
    "p-values: * < .05, ** < .01, *** < .001"
  )) |>
  group_tt(
    i = list(
      "General Stress ~ K Mean"               = 1,
      "Depression Score ~ K Mean"             = 3,
      "General Health ~ K Mean"               = 5,
      "Health Costs Due to Politics ~ K Mean" = 7
    )
  ) |>
  style_tt(i = c(1, 4, 7, 10), bold = TRUE)

s6_table

s6_table |> save_tt(here(table_dir, "s6_health_correlates.docx"), overwrite = TRUE)
```

## Bayes Corrs

```{r}
focal_outcomes   <- c("k_all", "k_outgroup_ingroup", "k_ingroup", "k_outgroup", "k_mean")
focal_predictors <- c("democratic_support", "vote_importance", "party_identity_strength", "affpol_weighted",
                      "stress_general", "depression", "health_general", "health_politics")

predictor_labels <- c(
  "democratic_support"      = "Support for Democracy",
  "vote_importance"         = "Voting Importance",
  "party_identity_strength" = "Partisan Identity Strength",
  "affpol_weighted"         = "Affective Polarization",
  "stress_general"          = "General Stress",
  "depression"              = "Depression Score",
  "health_general"          = "General Health",
  "health_politics"         = "Health Costs Due to Politics"
)

outcome_labels <- c(
  "k_all"              = "K All",
  "k_outgroup_ingroup" = "K Out\u2212In",
  "k_ingroup"          = "K In",
  "k_outgroup"         = "K Out",
  "k_mean"             = "K Mean"
)

bayes_table <- model_data |>
  select(all_of(c(focal_outcomes, focal_predictors))) |>
  correlation(bayesian = TRUE) |>
  as_tibble() |>
  filter(Parameter1 %in% focal_outcomes, Parameter2 %in% focal_predictors) |>
  mutate(
    bf   = if_else(BF > 1000, "> 1000", sprintf("%.2f", BF)),
    cell = paste0(sprintf("%.2f", rho), " [BF\u2081\u2080 = ", bf, "]"),
    Predictor = factor(recode(Parameter2, !!!predictor_labels), levels = unname(predictor_labels)),
    Outcome   = factor(recode(Parameter1, !!!outcome_labels),   levels = unname(outcome_labels))
  ) |>
  select(Predictor, Outcome, cell) |>
  arrange(Predictor, Outcome) |>
  pivot_wider(names_from = Outcome, values_from = cell) |>
  tt(notes = "Cells show the posterior median correlation (r) followed by the Bayes factor in favor of a nonzero correlation (BF\u2081\u2080).")

bayes_table

bayes_table |> save_tt(here(table_dir, "s_bayes_correlations.docx"), overwrite = TRUE)
```

```{r}
#| fig-width: 10
#| fig-height: 7
#| fig-dpi: 600

bayes_plot_data <- model_data |>
  select(all_of(c(focal_outcomes, focal_predictors))) |>
  correlation(bayesian = TRUE) |>
  as_tibble() |>
  filter(Parameter1 %in% focal_outcomes, Parameter2 %in% focal_predictors) |>
  mutate(
    bf_label  = if_else(BF > 1000, "> 1000", sprintf("%.2f", BF)),
    label = paste0(sprintf("%.2f", rho), "\nBF10 = ", bf_label),
    Predictor = factor(recode(Parameter2, !!!predictor_labels), levels = rev(unname(predictor_labels))),
    Outcome   = factor(recode(Parameter1, !!!outcome_labels),   levels = unname(outcome_labels))
  )

bayes_corr_plot <- bayes_plot_data |>
  ggplot(aes(x = Outcome, y = Predictor, fill = rho)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = label), size = 4.5, lineheight = 0.9, family = "sans") +
  scale_fill_gradient2(
    low      = "#2c7bb6",
    mid      = "white",
    high     = "#d7191c",
    midpoint = 0,
    limits   = c(-0.3, 0.3),
    oob      = scales::squish,
    name     = "r"
  ) +
  scale_x_discrete(position = "top") +
  labs(x = NULL, y = NULL) +
  theme_nice() +
  theme(
    panel.grid       = element_blank(),
    axis.text.x.top  = element_text(angle = 0, hjust = 0.5),
    legend.position  = "right"
  )

bayes_corr_plot

ggsave(
  here(fig_dir, "bayes_correlation_heatmap.png"),
  plot   = bayes_corr_plot,
  height = 7,
  width  = 10,
  dpi    = 600
)

```


